---
title: "Yield loss data analysis"
output: html
---


Setup chunk: Load data and libraries. 

```{r, Setup, echo = FALSE, message = FALSE}
## libraries #####

library(tidyverse)
library(lme4)
library(mgcv)
library(itsadug)
library(reghelper)
library(wesanderson)

# data #####
all_data <- readRDS("../../data/all_data_2020.08.07.rds")
all.data.stan <- readRDS("../../data/all_data_stan_2020.08.07.rds")

```

To start we'll do a series of simple exploratory plots to see whether or not we choose to use soil data summarized using mean or geometric mean, and whether or not we choose to use SOC or SOM as our metric of choice. 

```{r, Plot yield by SOC}

ggplot(data = all_data, aes(x = ssurgo_clay_mean, y = ssurgo_h_square))+
  geom_jitter()+
  geom_smooth()

ggplot(data = all_data, aes(x = ssurgo_h, y = Yield_decomp_mult))+
  geom_jitter()+
  geom_smooth()

ggplot(data = all_data, aes(x = ssurgo_om_mean, y = Yield_decomp_mult))+
  geom_jitter()+
  geom_smooth()

ggplot(data = all_data, aes(x = ssurgo_clay_mean, y = Yield_decomp_mult))+
  geom_jitter()+
  geom_smooth()
```
Next, we'll develop simple linear mixed models to analyze the data. 

```{r}
full.main <- all.data.stan %>%
  lmerTest::lmer(
    data = .,
    formula = Yield_decomp_mult ~ summer_spei + ssurgo_om_mean + ssurgo_clay_mean +
      ssurgo_h + ssurgo_cec_mean + (1 | state_alpha)
  )

car::vif(full.main)
summary(full.main)

```
VIF for CEC is very high. Remove CEC and refit. 

```{r,}

full.main <- all.data.stan %>%
  lmerTest::lmer(data = ., 
           formula = Yield_decomp_mult ~ summer_spei+ssurgo_om_mean+ssurgo_clay_mean+ssurgo_h+(1|state_alpha))


car::vif(full.main)
summary(full.main)

```

VIFs are low now. Final variables will be SOM, Clay, and H+ concentration. 



```{r, SOM:SPEI 2-way interaction}

interactions::interact_plot(
  model = m.county,
  pred = ssurgo_om_mean,
  modx = summer_spei,
  modx.values = quantile(all.data.stan$summer_spei),
  partial.residuals = T
)
```


As SPEI decreases, drought severity increases, SOM effect is stronger. 

```{r, SOM:SPEI:clay 3-way interaction}

interactions::interact_plot(
  model = m.county,
  pred = ssurgo_om_mean,
  mod2 = summer_spei,
  mod2.values = quantile(all.data.stan$summer_spei)[2:4],
  modx = ssurgo_clay_mean,
  modx.values = quantile(all.data.stan$ssurgo_clay_mean)[2:4],
  partial.residuals = F
)
```
Clay:SOM:SPEI. Yields increase with SOM more strongly under drought. Clay amplifies this effect. This appears to be largely because of counties that are high clay, low SOM. 

```{r, SOM:SPEI:H+ 3-way interaction}

interactions::interact_plot(
  model = m.county,
  pred = ssurgo_om_mean,
  mod2 = summer_spei,
  mod2.values = quantile(all.data.stan$summer_spei)[1:3],
  modx = ssurgo_h,
  modx.values = quantile(all.data.stan$ssurgo_h)[1:3],
  partial.residuals = F
)
```

```{r, Clay:SPEI:H+ 3-way interaction}

interactions::interact_plot(
  model = m.county,
  pred = ssurgo_clay_mean,
  mod2 = summer_spei,
  mod2.values = quantile(all.data.stan$summer_spei)[1:3],
  modx = ssurgo_h,
  modx.values = quantile(all.data.stan$ssurgo_h)[1:3],
  partial.residuals = F
)
```

```{r}
mean(all_data$ssurgo_om_mean)

test <- filter(all_data,ssurgo_om_mean < 2.9, spei.cut != "Normal")

quantile(test$Yield_decomp_mult)

sd(test$Yield_decomp_mult)/mean(test$Yield_decomp_mult)

```

```{r,}
# All conditions

all.data.stan %>%
  filter(spei.cut %in% c("Very severe", "Severe", "Moderate")) %>%
  ggplot(data = ., aes(x = ssurgo_om_mean, y = Yield_decomp_add))+ 
  geom_point()

all.conditions <- all.data.stan %>%
  lm(data = ., formula = Yield_decomp_mult ~ ssurgo_om_mean*ssurgo_clay_mean*ssurgo_h)


plot(all.conditions)
summary(all.conditions)


mean(all_data$summer_spei)-2*sd(all_data$summer_spei)
```


```{r,}
# All drought, SPEI less than mean

all.data.stan %>%
  filter(spei.cut %in% c("Very severe", "Severe", "Moderate")) %>%
  ggplot(data = ., aes(x = ssurgo_om_mean, y = Yield_decomp_mult))+ 
  geom_point()

all.drought <- all.data.stan %>%
  filter(spei.cut %in% c("Very severe", "Severe", "Moderate")) %>%
  lmerTest::lmer(data = ., formula = Yield_decomp_mult ~ ssurgo_om_mean*ssurgo_clay_mean*ssurgo_h+(1|state_alpha))


plot(all.drought)
summary(all.drought)



```

```{r,}

# Very severe drought

all_data %>%
  filter(spei.cut == "Very severe") %>%
  ggplot(data = ., aes(x = ssurgo_h, y = Yield_decomp_mult)) + geom_point() 
  
very_svr <- all.data.stan %>%
  filter(spei.cut == "Very severe") %>%
  lmerTest::lmer(data = ., formula = Yield_decomp_mult ~ ssurgo_om_mean*ssurgo_clay_mean*ssurgo_h+(1|state_alpha))


summary(very_svr)


interactions::interact_plot(
  model = very_svr,
  pred = ssurgo_om_mean,
  modx = ssurgo_h,
  modx.values = c(quantile(all.data.stan$ssurgo_h)[2:3], 0.7),
  plot.points  = T
)


```

```{r,}
# severe drought
all.data.stan %>%
  filter(spei.cut == "Severe") %>%
  ggplot(data = ., aes(x = ssurgo_om_mean, y = Yield_decomp_mult)) + geom_point() 
  
severe <- all.data.stan %>%
  filter(spei.cut == "Severe") %>%
  lmerTest::lmer(data = ., formula = Yield_decomp_mult ~ ssurgo_om_mean*ssurgo_clay_mean*ssurgo_h+(1|state_alpha))

summary(severe)

```


```{r,}

# Moderate drought, SPEI > -0.4264771, < 0.1208768

all.data.stan %>%
  filter(spei.cut == "Moderate") %>%
  ggplot(data = ., aes(x = ssurgo_om_mean, y = Yield_decomp_mult)) + geom_point() 
  
moderate <- all.data.stan %>%
  filter(spei.cut == "Moderate") %>%
  lmerTest::lmer(data = ., formula = Yield_decomp_mult ~ ssurgo_om_mean*ssurgo_clay_mean*ssurgo_h+(1|state_alpha))

summary(moderate)

interactions::interact_plot(
  model = moderate,
  pred = ssurgo_om_mean,
  modx = ssurgo_h,
  modx.values = quantile(all.data.stan$ssurgo_h)[3:5],
  mod2 = ssurgo_clay_mean,
  mod2.values = quantile(all.data.stan$ssurgo_clay_mean)[1:5],
  plot.points  = T
)
```

```{r,normal to above average precip}

all.data.stan %>%
  filter(spei.cut == "Normal") %>%
  ggplot(data = ., aes(x = ssurgo_om_mean, y = Yield_decomp_mult)) + geom_point() 
  
normal <- all.data.stan %>%
  filter(spei.cut == "Normal") %>%
  lmerTest::lmer(data = ., formula = Yield_decomp_mult ~ ssurgo_om_mean*ssurgo_clay_mean*ssurgo_h+(1|state_alpha))

summary(normal)


```
```{r}

mod <- function(df){
  lmerTest::lmer(data = df, Yield_decomp_mult ~ ssurgo_om_mean*ssurgo_clay_mean*ssurgo_h+(1|state_alpha))
}

model_by_drought <- all.data.stan %>%
  group_by(spei.cut) %>%
  nest() %>%
  mutate(model_by_drought = map(data, mod))

View(model_by_drought %>% 
  mutate(glance = map(model_by_drought, broom.mixed::tidy)) %>% 
  unnest(glance) %>%
  mutate(p.value = round(p.value, digits = 5)) %>%
  filter(effect == "fixed", 
         term != "(Intercept)") %>%
  arrange(spei.cut) %>%
  select(-data, -model_by_drought, -effect, - group))
```