---
title: "Crop insurance data analysis"
output: html
---

Setup chunk: Load data and libraries.

```{r, Setup, echo = FALSE, message = FALSE}
## libraries #####

library(tidyverse)
library(lme4)
library(mgcv)
library(itsadug)
library(reghelper)
library(tidyverse)
library(VGAM)
library(censReg)
library(stargazer)
library(sampleSelection)



# data #####
all_data <- readRDS("../../data/all_data_2020.08.07.rds")
all.data.stan <- readRDS("../../data/all_data_stan_2020.08.07.rds")

all.data.stan <- all.data.stan %>%
  filter(loss_cost < 300)


```

To start, we'll do a simple model of indemnities as a function of drought and various soil properties. 
Next, we'll develop simple linear mixed models to analyze the data. 

```{r}
full.main <- all.data.stan %>%
  lmerTest::lmer(
    data = .,
    formula = loss_cost ~ summer_spei + ssurgo_om_mean + ssurgo_clay_mean +
      ssurgo_h + ssurgo_cec_mean + (1 | GEOID)
  )

car::vif(full.main)
summary(full.main)

```
VIF for CEC is very high. Remove CEC and refit. 

```{r,}

full.main <- all.data.stan %>%
  lmerTest::lmer(data = ., 
           formula = loss_cost ~ summer_spei+ssurgo_om_mean+ssurgo_clay_mean+ssurgo_h+(1|GEOID))


car::vif(full.main)
summary(full.main)

plot(full.main)

```

In a simple model, drought (july_spei) emerges as significant negative coefficient for indemnities - as july spei goes up, conditions get less droughty, and insurance payments go down. Interestingly, an interaction effect between SOC and july_spei emerges but in a direction that contradicts our hypothesis - SOC interacts with drought in an additive manner, as drought gets worse payouts are higher when SOC is higher. However, this pattern may be explained by the fact that SOC has a positive effect on yields, as demonstrated in our previous analysis, meaning that counties with higher SOC have higher yield potential to begin with and are more likely to be insured under adverse conditions. 


Model coefficients change only very slightly, and liabilties emerge as a very strong effect (highest beta coefficient, least p value). 

However, given that there are a significant number of observations for which indemnities are zero or very close to zero, simple LMEs are likely not appropriate for these questions. Tobit regression is an approach for when the dependent variable is zero-inflated (or has a distinct lower limit after data are standardized) and continuous.

``` {r, tobit models}

m <- all.data.stan %>%
  vglm(data = ., formula =loss_cost ~ summer_spei*ssurgo_om_mean*ssurgo_clay_mean*ssurgo_h,
       tobit(Lower =  min(all.data.stan$loss_cost), type.fitted = "censored"), maxit = 50, smart = TRUE)

summary(m)


```

Using a tobit regression, the effects structure and size of coefficients remains largely the same. An interaction effect for SOC and drought emerges, and in terms of directionality it is similar to simpler models- SOC reduces payouts under adverse conditions. 

In addition to a tobit model, we'll run a Heckman selection model. A Heckman selection model is similar but rather than assuming that the process that creates "zeroes" in the dependent variable is latent, it assumes that "zero" observations of the dependent variable are based on some selection/threshold process that can be modeled by the data in hand. In this case, we'll model a selection process in which payouts are non-zero as a function of drought, and then any non-zero payouts are further predicted by a combination of drought and soil properties.


LOSS COST MODELS 

```{r,}
# All drought, SPEI less than mean

all.data.stan %>%
  filter(spei.cut %in% c("Very severe", "Severe", "Moderate")) %>%
  ggplot(data = ., aes(x = ssurgo_om_mean, y = loss_cost))+ 
  geom_point()

all.drought <- all.data.stan %>%
  mutate(non.zero = as.logical(.$loss_cost != 0)) %>%
  filter(spei.cut %in% c("Very severe", "Severe", "Moderate")) %>%
  #lm(data = ., formula = loss_cost ~ssurgo_om_mean*ssurgo_clay_mean*ssurgo_h) #%>%
  vglm(data = ., formula = loss_cost ~ ssurgo_om_mean*ssurgo_clay_mean*ssurgo_h,
       tobit(Lower = min(all.data.stan$loss_cost),Upper = 100,  type.fitted = "censored"), maxit = 50, smart = TRUE)

all.drought2 <-all.data.stan %>%
  mutate(non.zero = as.logical(.$loss_cost != 0)) %>%
  filter(spei.cut %in% c("Very severe", "Severe", "Moderate")) %>%
  lmerTest::lmer(data = ., formula = loss_cost ~ssurgo_om_mean*ssurgo_clay_mean*ssurgo_h+(1|GEOID))

summary(all.drought)

min(all_data$loss_cost)

# interactions::interact_plot(
#   model = all.drought,
#   pred = ssurgo_om_mean,
#   modx = ssurgo_clay_mean,
#   modx.values = quantile(all.data.stan$ssurgo_clay_mean),
#   mod2 = ssurgo_h,
#   mod2.values = quantile(all.data.stan$ssurgo_h),
#   partial.residuals = T
# )


all.data.stan %>%
  filter(ssurgo_h >1,
         ssurgo_clay_mean < -.3) %>%
  ggplot(data = ., aes(x = ssurgo_om_mean, y = loss_cost))+
  geom_point()

```

```{r,}
# Very severe drought
all.data.stan %>% 
  filter(spei.cut %in% c("Very severe"), ssurgo_clay_mean >= 1) %>%
  ggplot(data = ., aes(x = ssurgo_om_mean, y = loss_cost)) + 
  geom_point() 

vsvr.dat <- all.data.stan %>%
  filter(spei.cut == "Very severe") %>%
  #filter(!(ssurgo_h <= -0.25 & state_ansi == 19)) %>%
  plm::pdata.frame(x = ., c("state_alpha"))


test <- all_data %>%
  filter(spei.cut == "Very severe") %>%
  #filter(ssurgo_h <= quantile(.$ssurgo_h)[2]) %>%
  lmerTest::lmer(data = ., loss_cost ~ ssurgo_om_mean*ssurgo_clay_mean*ssurgo_h+(1|state_alpha)) 

  
summary(test)

# interactions::interact_plot(
#   model = test,
#   pred = ssurgo_om_mean,
#   modx = ssurgo_clay_mean,
#   modx.values = c(-0.5, 0, 0.5),
#   mod2 = ssurgo_h,
#   mod2.values = c( 0, 0.5, 1),
#   plot.points = T
# )



very_svr <- vsvr.dat %>%
  censReg( loss_cost ~ ssurgo_om_mean*ssurgo_clay_mean*ssurgo_h, left = 0, data = .)

summary(very_svr)


```

```{r,}
# severe drought
all.data.stan %>%
  filter(spei.cut == "Severe") %>%
  ggplot(data = ., aes(x = ssurgo_h, y = loss_cost)) + geom_point() 
  
severe.dat <- all.data.stan %>%
  filter(spei.cut == "Severe") %>%
  #filter(!(ssurgo_h>0 & ssurgo_om_mean > 1 & loss_cost == 0)) %>%
  plm::pdata.frame(x = ., c("state_alpha"))

  
severe <- severe.dat %>%
  censReg( loss_cost ~ ssurgo_om_mean*ssurgo_clay_mean*ssurgo_h, data = .)

summary(severe)

test <- all.data.stan %>%
  filter(spei.cut == "Severe", !(ssurgo_h>0 & ssurgo_om_mean > 0)) %>%
  lmerTest::lmer(data = ., loss_cost ~ ssurgo_om_mean*ssurgo_clay_mean*ssurgo_h+(1|state_alpha)) 

summary(test)

interactions::interact_plot(
  model = test,
  pred = ssurgo_om_mean,
  modx = ssurgo_h,
  modx.values = quantile(all.data.stan$ssurgo_h),
  # mod2 = ssurgo_h,
  # mod2.values = quantile(all.data.stan$ssurgo_h),
  plot.points = T
)

```


```{r,}

# Moderate drought, SPEI > -0.4264771, < 0.1208768

all.data.stan %>%
  filter(spei.cut == "Moderate") %>%
  ggplot(data = ., aes(x = ssurgo_om_mean, y = loss_cost)) + geom_point() 

moderate.dat <- all.data.stan %>%
  filter(spei.cut == "Moderate") %>%
  #filter(!(ssurgo_h < -0.25 & ssurgo_om_mean < -0.36 )) %>%
  plm::pdata.frame(x = ., c("state_alpha"))

(all.data.stan %>%
  filter(spei.cut == "Moderate", (ssurgo_h < -0.25 & ssurgo_om_mean < -0.36 )) %>%
  plm::pdata.frame(x = ., c("state_alpha")))
  
moderate <- moderate.dat %>%
  censReg( loss_cost ~ ssurgo_om_mean*ssurgo_clay_mean*ssurgo_h, data = .)
  
summary(moderate)


test <- all.data.stan %>%
  filter(spei.cut == "Moderate", ssurgo_h > -0.25) %>%
  lmerTest::lmer(data = ., loss_cost ~ ssurgo_om_mean*ssurgo_clay_mean*ssurgo_h+(1|state_alpha)) 

summary(test)

interactions::interact_plot(
  model = test,
  pred = ssurgo_om_mean,
  modx = ssurgo_h,
  modx.values = c(0, 0.5,1),
  # mod2 = ssurgo_h,
  #mod2.values = quantile(all.data.stan$ssurgo_om_mean),
  plot.points = T
)
3977-3948



```

```{r,normal to above average precip}

all.data.stan %>%
  filter(spei.cut == "Normal") %>%
  ggplot(data = ., aes(x = ssurgo_om_mean, y = Yield_decomp_mult)) + geom_point() 
  
normal.dat <- all.data.stan %>%
  filter(spei.cut == "Normal") %>%
  plm::pdata.frame(x = ., c("state_alpha"))

normal <- normal.dat %>%
  censReg( loss_cost ~ ssurgo_om_mean*ssurgo_clay_mean*ssurgo_h, data = .)



```


```{r}

vs_0 <- vsvr.dat %>%
  censReg( loss_cost ~ 1, data = .)
1 - as.vector(logLik(very_svr)/logLik(vs_0))

s_0 <- severe.dat %>%
  censReg( loss_cost ~ 1, data = .)
1 - as.vector(logLik(severe)/logLik(s_0))

m_0 <- moderate.dat %>%
  censReg( loss_cost ~ 1, data = .)
1 - as.vector(logLik(moderate)/logLik(m_0))

n_0 <- normal.dat %>%
  censReg( loss_cost ~ 1, data = .)
1 - as.vector(logLik(normal)/logLik(n_0))


```

CONCLUSIONS:

Using multiple different modeling approaches that account for the fact that payouts are first and foremost a function of the amount of coverage in dollars in a given county, we demonstrate that SOC has a mitigating impact on payouts under drought conditions. In some modeling approaches this impact is also conditional on native clay content such that the impact is greater on soils with lower clay.   

```{r}

View(bind_rows(data.frame(row.names(summary(very_svr)$estimate),summary(very_svr)$estimate), 
          data.frame(summary(severe)$estimate),
          data.frame(summary(moderate)$estimate),
          data.frame(summary(normal)$estimate)) %>%
  mutate(Pr...t. = round(Pr...t., digits = 5)) )
  





```



```{r}

rma <- all_data %>%
  mutate(speifct = cut(july_spei, breaks = c(-3.1,-2,-1,0,3.1), labels = c("Very severe", "Severe", "Moderate", "Normal")),
         soc.mg_ha = ssurgo_om_gm_mean,
         Corn.belt = case_when(Corn.belt == "CB"~"Corn belt",
                               Corn.belt == "NCB"~"Not corn belt")) %>%
  filter(Corn.belt == "Corn belt") %>%
  mutate(speifct=factor(speifct,levels=c("Normal","Moderate","Severe","Very severe"))) %>%
  arrange(speifct) %>%
  ggplot(data = .) + 
  scale_color_manual(values = c("#49A942",#"#A0C04D",
                                "#23487A",#"#0096D6",
                                "#90214A",#"#C5351C",
                                "#F3901D"))+
  geom_jitter(aes(x = soc.mg_ha, y = loss_cost, color = ssurgo_order_mode), alpha = 0.5, size = 2)+
  geom_smooth(aes(x = soc.mg_ha, y = loss_cost), method = "lm", color = "black", alpha = 0.2)+
  geom_abline(slope = 0, color = "dark grey") +
  facet_wrap(facets = "speifct", ncol = 4)+
  ylim(0,100)+
  labs(x = "SOC (Mg per ha)", y = "Proportion of liabilities paid (%)")+
  theme(legend.title = element_blank(), legend.position = "bottom")+
  theme(axis.title = element_text(size=12), axis.text.y = element_text(size=10),
          axis.text.x = element_text(size=10), legend.text = element_text(size = 10))

ggsave(rma, filename = "~/rma.jpg", width = 8, height = 4, units = "in")
```



```{r}

rma.svr <- all_data %>%
  mutate(speifct = cut(july_spei, breaks = c(-3.1,-2,-1,0,3.1), labels = c("Very severe", "Severe", "Moderate", "Normal")),
         soc.mg_ha = ssurgo_om_gm_mean,
         Corn.belt = case_when(Corn.belt == "CB"~"Corn belt",
                               Corn.belt == "NCB"~"Not corn belt")) %>%
  mutate(speifct=factor(speifct,levels=c("Normal","Moderate","Severe","Very severe"))) %>%
  arrange(speifct) %>%
  filter(speifct == c("Severe", "Very severe")) %>%
  ggplot(data = .) + 
  scale_color_manual(values = c("#49A942",#"#A0C04D",
                                "#23487A",#"#0096D6",
                                "#90214A",#"#C5351C",
                                "#F3901D"))+
  geom_jitter(aes(x = soc.mg_ha, y = loss_cost, color = ssurgo_order_mode), alpha = 0.5, size = 2)+
  geom_smooth(aes(x = soc.mg_ha, y = loss_cost), method = "lm", color = "black", alpha = 0.2)+
  geom_abline(slope = 0, color = "dark grey") +
  facet_grid(Corn.belt ~ speifct, )+
  labs(x = "SOM (%)", y = "Proportion of liabilities paid (%)")+
  theme(legend.title = element_blank(), legend.position = "bottom")+
  theme(axis.title = element_text(size=12), axis.text.y = element_text(size=10),
          axis.text.x = element_text(size=10), legend.text = element_text(size = 10))

ggsave(rma.svr, filename = "~/rma.svr.jpg", width = 6, height = 6, units = "in")
```




```{r}

# coeff

tribble(
  ~Drought, ~Coeff, ~std.error, ~p.value,
  "Normal", summary(normal)$coefficients[2,1], summary(normal)$coefficients[2,2], summary(normal)$coefficients[2,5],
  "Moderate",summary(moderate)$coefficients[2,1], summary(moderate)$coefficients[2,2], summary(moderate)$coefficients[2,5],
  "Severe",summary(severe)$coefficients[2,1], summary(severe)$coefficients[2,2], summary(severe)$coefficients[2,5],
  "Very severe",summary(very_svr)$coefficients[2,1], summary(very_svr)$coefficients[2,2], summary(very_svr)$coefficients[2,5],
) %>%
  mutate(Drought=factor(Drought,levels=c("Normal","Moderate","Severe","Very severe")))


```


```{r}

tribble(
  ~Drought, ~Coeff, ~std.error, ~p.value,
  "Normal", summary(normal)$coefficients[3,1], summary(normal)$coefficients[3,2], summary(normal)$coefficients[3,5],
  "Moderate",summary(moderate)$coefficients[3,1], summary(moderate)$coefficients[3,2], summary(moderate)$coefficients[3,5],
  "Severe",summary(severe)$coefficients[3,1], summary(severe)$coefficients[3,2], summary(severe)$coefficients[3,5],
  "Very severe",summary(very_svr)$coefficients[3,1], summary(very_svr)$coefficients[3,2], summary(very_svr)$coefficients[3,5],
) %>%
  mutate(Drought=factor(Drought,levels=c("Normal","Moderate","Severe","Very severe")))


```


```{r}

tribble(
  ~Drought, ~Coeff, ~std.error, ~p.value,
  "Normal", summary(normal)$coefficients[4,1], summary(normal)$coefficients[4,2], summary(normal)$coefficients[4,5],
  "Moderate",summary(moderate)$coefficients[4,1], summary(moderate)$coefficients[4,2], summary(moderate)$coefficients[4,5],
  "Severe",summary(severe)$coefficients[4,1], summary(severe)$coefficients[4,2], summary(severe)$coefficients[4,5],
  "Very severe",summary(very_svr)$coefficients[4,1], summary(very_svr)$coefficients[4,2], summary(very_svr)$coefficients[4,5],
) %>%
  mutate(Drought=factor(Drought,levels=c("Normal","Moderate","Severe","Very severe")))

```