---
title: "Yield data analysis"
output: html
---

Setup chunk: Load data and libraries. 

```{r, Setup, echo = FALSE, message = FALSE}
## libraries #####

library(tidyverse)
library(lme4)
library(mgcv)
library(itsadug)
library(reghelper)
library(wesanderson)

# data #####
all_data <- readRDS("../../data/all_data_2020.08.07.rds")
all.data.stan <- readRDS("../../data/all_data_stan_2020.08.07.rds")

```

To start we'll do a series of simple exploratory plots to see whether or not we choose to use soil data summarized using mean or geometric mean, and whether or not we choose to use SOC or SOM as our metric of choice. Yield_decomp_add is the de-trended yield data. I don't see geometric means in the data, data-processing code, or code below?

```{r, Plot yield by SOC}

ggplot(data = all_data, aes(x = ssurgo_clay_mean, y = ssurgo_h_square))+
  geom_jitter()+
  geom_smooth()

ggplot(data = all_data, aes(x = ssurgo_h, y = Yield_decomp_add))+
  geom_jitter()+
  geom_smooth()

ggplot(data = all_data, aes(x = ssurgo_om_mean, y = Yield_decomp_add))+
  geom_jitter()+
  # geom_smooth()   #looks like a saturation curve y=(x)^(1/2)
  geom_smooth(method="lm")
  
ggplot(data = all_data, aes(x = ssurgo_clay_mean, y = Yield_decomp_add))+
  geom_jitter()+
  geom_smooth()  # looks like it peaks then comes back down, like a normal curve (probability density function)

```
Next, we'll develop simple linear mixed models to analyze the data. 
BM: Why are we using linear models when the third plot above is exponential?

```{r}
full.main <- all.data.stan %>%
  lmerTest::lmer(
    data = .,
    formula = Yield_decomp_add ~ summer_spei + ssurgo_om_mean + ssurgo_clay_mean +
      ssurgo_h + ssurgo_cec_mean + (1 | GEOID)
  )

car::vif(full.main)
summary(full.main)

```
VIF for CEC is very high. Remove CEC and refit. 

```{r,}

full.main <- all.data.stan %>%
  lmerTest::lmer(data = ., 
           formula = Yield_decomp_add ~ summer_spei+ssurgo_om_mean+ssurgo_clay_mean+ssurgo_h+(1|GEOID))


car::vif(full.main)
summary(full.main)

```

VIFs are low now. Final variables will be SOM, Clay, and H+ concentration.

BM: GEOID is the county FIPS code and the random effect used by Dan in the above models. For P-SHEC we want to get to run regression on a sub-county spatial scale so that it is more relevant to an individual farm. However notice in the models below he uses state as the fixed effect rather than county, and state is what he reports using in the paper. How does the final model look if we go back to county?

```{r, Yield models with SPEI, echo = TRUE}

# SPEI as drought indicator

m.state <- all.data.stan %>%
  lmerTest::lmer(data = ., 
           formula = Yield_decomp_add ~ summer_spei*ssurgo_om_mean*ssurgo_h*ssurgo_clay_mean+(1|state_alpha))

summary(m.state)

ranef(m.state) # add ranef for state to intercept fixed effect below to get estimate for state
fixef(m.state)

broom.mixed::tidy(m.state) %>%
  mutate(p.value = round(p.value, digits = 5))

# Plot residuals and inspect
ggplot(data.frame(eta=predict(m.state,type="link"),pearson=residuals(m.state,type="pearson")),
       aes(x=eta,y=pearson)) +
  geom_point() +
  theme_bw()

```

Log likelihood ratio tests, anova, and conditional r-squared estimates indicate that the random effect of county improves model fit. 

BM: However he uses state. --in paper they argue that "since farms in same state are generally more likely to implement fert rates and management practices simimlar to other farms in their state than those out of their state, our model accounts to some degree for broad differences in management." This is because they were unable to account for differences in management in the data (not available for fertilizer rate, and tillage data only available recently in AgCensus, I'll also add,that not all respondents to the survey report tillage practices.)
Also where is the code showing these tests? Need a null model without county to compare. Below we run the same code chunk as above with GEOID (county) rather than state. And make a null model to compare.

```{r, Yield models with SPEI, echo = TRUE}

# SPEI as drought indicator

m.county <- all.data.stan %>%
  lmerTest::lmer(data = ., 
           formula = Yield_decomp_add ~ summer_spei*ssurgo_om_mean*ssurgo_h*ssurgo_clay_mean+(1|GEOID))

summary(m.county)

broom.mixed::tidy(m.county) %>%
  mutate(p.value = round(p.value, digits = 5))

# Plot residuals and inspect
ggplot(data.frame(eta=predict(m.county,type="link"), # link means predictions will be on the scale of the linear predictor rather than the response variable ("response" is the other option for type=)
                  pearson=residuals(m.county,type="pearson")),
       aes(x=eta,y=pearson)) +
  geom_point() +
  theme_bw()

# Null model
m.county.h0 <- all.data.stan %>%
  lm(data = ., 
           formula = Yield_decomp_add ~ summer_spei*ssurgo_om_mean*ssurgo_h*ssurgo_clay_mean)

# Compare null model and model with random effect of county
anova(m.county,m.county.h0)

# compare conditional R-squared estimates 
# about conditional vs. marginal R2 for mixed effects models see: https://jonlefcheck.net/2013/03/13/r2-for-linear-mixed-effects-models/
MuMIn::r.squaredGLMM(m.county)
MuMIn::r.squaredGLMM(m.county.h0)

```
BM: m.county model is significantly more parsimonious than the null model (both AIC and BIC are lower); has a higher log likelihood (better); residual deviance is lower; and the R2c (conditional R-squared) is higher (better fit). So, yes, county improves model fit. 

Below we'll look at the coefficients. 

```{r}

summary(m.state)

```
There are just a few interaction effects with SPEI. We'll evaluate them here. 

```{r, SOM:SPEI 2-way interaction}

interactions::interact_plot(
  model = m.state,
  pred = ssurgo_om_mean,
  modx = summer_spei,
  modx.values = quantile(all.data.stan$summer_spei),
  partial.residuals = T
)
```


As SPEI decreases, drought severity increases, SOM effect is stronger. 

```{r, SOM:SPEI:clay 3-way interaction}

interactions::interact_plot(
  model = m.state,
  pred = ssurgo_om_mean,
  mod2 = summer_spei,
  mod2.values = quantile(all.data.stan$summer_spei),
  modx = ssurgo_clay_mean,
  modx.values = quantile(all.data.stan$ssurgo_clay_mean),
  partial.residuals = T
)
```
Clay:SOM:SPEI. Yields increase with SOM more strongly under drought. Clay amplifies this effect. This appears to be largely because of counties that are high clay, low SOM. 

```{r, SOM:SPEI:H+ 3-way interaction}

interactions::interact_plot(
  model = m.state,
  pred = ssurgo_om_mean,
  mod2 = summer_spei,
  mod2.values = quantile(all.data.stan$summer_spei),
  modx = ssurgo_h,
  modx.values = quantile(all.data.stan$ssurgo_h)[3:5],
  partial.residuals = T
)
```
OM:SPEI:H+. Yields increase with SOM more strongly under drought. As drought severity decreases, SOM has a more strongly negative effect on yields when soils are also very high H+ concentrations. This also seems to be due to very few points that are high pH.

Interaction plots indicate that SOM interacts with SPEI to mitigate the effect of low SPEI (drought) on yields. SOM has a greater positive effect on yields under all conditions when clay content is high. Higher clay content suppresses yields under drought conditions. There are additional interaction effects with SOM and SPEI for H+ and clay. The interaction with clay simply enhances the same effect, whereas the one with H+ reverses the effect at very high H+ concentrations. 

Now we'll check if the same patterns hold when we use DSCI as the drought indicator instead of SPEI. 

```{r, Yield models with SOC and DSCI, echo = TRUE}
# DSCI as drought indicator

m.state.dsci <- all.data.stan %>%
  lmerTest::lmer(data = ., 
           formula = Yield_decomp_add ~ DSCI.mean*ssurgo_om_mean*ssurgo_h*ssurgo_clay_mean+(1|state_alpha))

# View(broom.mixed::tidy(m.state.dsci) %>%
#   mutate(p.value = round(p.value, digits = 5)))

```

They mostly agree. Directionality of coeffecients is the same, as are significance factors and size of coefficients. An additional three-way interaction with DSCI:CLAY:H+ emerges. Since that does not affect the main hypotheses, we're not concerned. Given agreement between DSCI and SPEI models, we'll elect to use SPEI going forward as it introduces fewer possible artefacts from observer bias and is more strongly tied to physcial metrics.

Next, we'll generate similar plots on the subset of observations below 0 SPEI given the interesting patterns revealed by the above figures. 
```{r,}
# All conditions no SPEI 

all.data.stan %>%
  ggplot(data = ., aes(x = ssurgo_om_mean, y = Yield_decomp_add))+ 
  geom_point()

all.conditions <- all.data.stan %>%
  lmerTest::lmer(data = ., formula = Yield_decomp_add ~ ssurgo_om_mean*ssurgo_clay_mean*ssurgo_h+(1|state_alpha))

plot(all.conditions)
summary(all.conditions)

```


```{r,}
# All drought, SPEI less than mean

all.data.stan %>%
  filter(spei.cut %in% c("Very severe", "Severe", "Moderate")) %>%
  ggplot(data = ., aes(x = ssurgo_om_mean, y = Yield_decomp_add))+ 
  geom_point()

all.drought <- all.data.stan %>%
  filter(spei.cut %in% c("Very severe", "Severe", "Moderate")) %>%
  lmerTest::lmer(data = ., formula = Yield_decomp_add ~ ssurgo_om_mean*ssurgo_clay_mean*ssurgo_h+(1|state_alpha))


plot(all.drought)
summary(all.drought)



```

```{r,}
# Very severe drought

all.data.stan %>%
  filter(spei.cut == "Very severe") %>%
  ggplot(data = ., aes(x = ssurgo_h, y = Yield_decomp_add)) + geom_point() 
  
very_svr <- all.data.stan %>%
  filter(spei.cut == "Very severe") %>% 
  lmerTest::lmer(data = ., formula = Yield_decomp_add ~ ssurgo_om_mean*ssurgo_clay_mean*ssurgo_h+(1|state_alpha))

summary(very_svr)
MuMIn::r.squaredGLMM(very_svr)


```

```{r,}
# severe drought
all.data.stan %>%
  filter(spei.cut == "Severe") %>%
  ggplot(data = ., aes(x = ssurgo_om_mean, y = Yield_decomp_add)) + geom_point() 
  
severe <- all.data.stan %>%
  filter(spei.cut == "Severe",
         # !(ssurgo_clay_mean > 0 & ssurgo_om_mean > 1),
         # !(ssurgo_clay_mean <= -.25 & ssurgo_om_mean > 1)
         !(ssurgo_h > 0 & ssurgo_om_mean >1)
         ) %>%
  lmerTest::lmer(
    data = .,
    formula = Yield_decomp_add ~ ssurgo_om_mean * ssurgo_clay_mean * ssurgo_h +
      (1 | state_alpha)
  )

summary(severe)
MuMIn::r.squaredGLMM(severe)
car::vif(severe)

1537

interactions::interact_plot(
  model = severe,
  pred = ssurgo_om_mean,
  modx = ssurgo_clay_mean,
  modx.values = c(-0.5,-0.25, 0, 0.25),
  mod2 = ssurgo_h,
  mod2.values = c(-0.5,-0.25, 0, 0.25,1),
  plot.points = T
)

```


```{r,}

# Moderate drought, SPEI > -0.4264771, < 0.1208768

all.data.stan %>%
  mutate(hfct = cut(ssurgo_h, breaks = quantile(all.data.stan$ssurgo_h))) %>%
  mutate(hfct = cut(ssurgo_om_mean, breaks = c(-3, -1, -0.5,0, 0.5, 1, 3.2))) %>%
  filter(spei.cut == "Moderate") %>%
  ggplot(data = ., aes(x = ssurgo_h, y = Yield_decomp_add)) + geom_point() + geom_smooth(method = "lm") + facet_wrap("hfct")
  
moderate <- all.data.stan %>%
  filter(spei.cut == "Moderate",
         !(ssurgo_h > 0 & ssurgo_om_mean >1)
         ) %>%
  lmerTest::lmer(data = ., formula = Yield_decomp_add ~ ssurgo_om_mean*ssurgo_clay_mean*ssurgo_h+(1|state_alpha))

summary(moderate)
MuMIn::r.squaredGLMM(moderate)

3998 -3956

interactions::interact_plot(
  model = moderate,
  pred = ssurgo_om_mean,
  modx = ssurgo_h,
  modx.values = quantile(all.data.stan$ssurgo_h),
plot.points = T)

```

```{r,normal to above average precip}

all.data.stan %>%
  filter(spei.cut == "Normal") %>%
  ggplot(data = ., aes(x = ssurgo_om_mean, y = Yield_decomp_add)) + geom_point() 
  
normal <- all.data.stan %>%
  filter(spei.cut == "Normal") %>%
  lmerTest::lmer(data = ., formula = Yield_decomp_add ~ ssurgo_om_mean*ssurgo_clay_mean*ssurgo_h+(1|state_alpha))

summary(normal)
MuMIn::r.squaredGLMM(normal)

interactions::interact_plot(
  model = normal,
  pred = ssurgo_h,
  modx = ssurgo_om_mean,
  modx.values = c(-0.5, 0, 0.5),
  mod2 = ssurgo_clay_mean,
  mod2.values = c(-0.5, 0, 0.5),
plot.points = T)


```
We ran models on different subsets of data to represent different levels of drought intensity. No matter the level of drought, SOC had a positive, significant effect on yields. At more intense levels of drought, that effect was more stronger, i.e. the coefficient was greater. Based on these data, we conclude that counties with higher SOC content per estimations in the gSSURGO database have greater resilience to yield losses in drought years.  


```{r}

mod <- function(df){
  lmerTest::lmer(data = df, Yield_decomp_add ~ ssurgo_om_mean*ssurgo_clay_mean*ssurgo_h+(1|state_alpha))
}



model_by_drought <- all.data.stan %>%
  group_by(spei.cut) %>%
  nest() %>%
  mutate(model_by_drought = map(data, mod))

# this table contains the slopes used in Fig. 1, caption reads "trendlines represent
# predicted yields based on that marginal effect [of SOM]."
coeff.table <- model_by_drought %>% 
  mutate(glance = map(model_by_drought, broom.mixed::tidy)) %>% 
  unnest(glance) %>%
  mutate(p.value = round(p.value, digits = 5)) %>%
  filter(effect == "fixed", 
         term != "(Intercept)") %>%
  arrange(spei.cut) %>%
  select(-data, -model_by_drought, -effect, - group)
```



```{r, Presentation figures}


test <-   # Create a new dataframe to add predictions to
  plyr::ldply(unique(all.data.stan$state_alpha), function(x) {
    data.frame(
      state_alpha = x,
      ssurgo_om_mean = seq(-3, 4, 0.05),
      ssurgo_clay_mean = mean(all.data.stan$ssurgo_clay_mean),
      ssurgo_h = mean(all.data.stan$ssurgo_h)
    )})
   
som.rev <- function(x){
  (x*(2*sd(all_data$ssurgo_om_mean)))+mean(all_data$ssurgo_om_mean)
}  # this reverses the standardized SOM data, the formula for that was (x-mean(x))/(2*sd(x))

test2 <- test %>%  # Add predictions, these are used for fitting the smoothed line in the ggplot below
  mutate(Normal = predict(object = normal, newdata = .),  # normal is an object in the environment, formal class lmerModLmerTest; The prediction intervals are for a single observation at each case in newdata (or by default, the data used for the fit)
         Moderate = predict(object = moderate, newdata = .),
         Severe = predict(object = severe, newdata = .),
         "Very severe" = predict(object = very_svr, newdata = .),
         ssurgo_om_mean = som.rev(ssurgo_om_mean)) %>%
  filter(ssurgo_om_mean > 0, ssurgo_om_mean < 10) %>%
  select(ssurgo_om_mean, Normal, Moderate, Severe, 'Very severe') %>%
  pivot_longer(-ssurgo_om_mean, names_to = "spei.cut", values_to = "predicted.yield")


test2$spei.cut_f = factor(test2$spei.cut, levels=c("Normal","Moderate","Severe","Very severe"))

# ggplot(test2, aes(x=om_descale, y=pred))+
#   geom_smooth(method = "lm")


all_data$spei.cut_f = factor(all_data$spei.cut, levels=c("Normal","Moderate","Severe","Very severe"))



yield.plot <- all_data %>%
  mutate(ssurgo_order_mode = factor(
    ssurgo_order_mode,
    levels = c(
      "Alfisol",
      "Mollisols",
      "Entisols",
      "Inceptisols",
      "Spodosols",
      "Ultisols",
      "Vertisols"
    )
  )) %>%
  # mutate(spei.cut=factor(spei.cut,levels=c("Normal","Moderate","Severe","Very severe"))) %>%
  # arrange(spei.cut) %>%
  ggplot(data = .) +
  scale_color_manual(values = c(
    "#004949",
    "#009292",
    "#ff6db6",
    "#ffb6db",
    "#490092",
    "#006ddb",
    "#b66dff"
  )) +
  geom_jitter(
    aes(x = ssurgo_om_mean, y = Yield_decomp_add, color = ssurgo_order_mode),
    alpha = 0.4,
    size = 1
  ) +
  geom_smooth(
    data = test2,
    aes(x = ssurgo_om_mean, y = predicted.yield),
    method = "lm",
    color = "black"
  ) +
  facet_grid(. ~ spei.cut_f) +
  labs(x = "SOM (%)", y = expression(Yield ~ Mg ~ ha ^ -1)) +
  ylim(0, 15) +
  theme(legend.title = element_blank(), legend.position = "bottom") +
  theme(
    axis.title = element_text(size = 12),
    axis.text.y = element_text(size = 9),
    axis.text.x = element_text(
      size = 9,
      angle = 45,
      hjust = 1
    ),
    legend.text = element_text(size = 10)
  ) + 
  guides(colour = guide_legend(override.aes = list(size=5, alpha = 0.8)))

dat_text <- data.frame(
  spei.cut = c("Normal","Moderate","Severe","Very severe"),
  slope = c(0.73, 0.76, 1.18, 2.2)  # these come from coeff.table
)

dat_text$spei.cut_f = factor(dat_text$spei.cut, levels=c("Normal","Moderate","Severe","Very severe"))


yield.plot <- yield.plot + geom_text(
  size = 3.25,
  data    = dat_text,
  mapping = aes(x = 9.5, y = 14.25, label = slope),
  hjust   = "right",
  vjust   = "top"
)

ggsave(yield.plot, filename = "~/yield.jpg", width = 8, height = 4, units = "in")


```
```{r}

yield_loss_plot <- all_data %>%
  mutate(ssurgo_order_mode = factor(
    ssurgo_order_mode,
    levels = c(
      "Alfisol",
      "Mollisols",
      "Entisols",
      "Inceptisols",
      "Spodosols",
      "Ultisols",
      "Vertisols"
    )
  )) %>%
  # mutate(spei.cut=factor(spei.cut,levels=c("Normal","Moderate","Severe","Very severe"))) %>%
  # arrange(spei.cut) %>%
  ggplot(data = .) +
  scale_color_manual(values = c(
    "#004949",
    "#009292",
    "#ff6db6",
    "#ffb6db",
    "#490092",
    "#006ddb",
    "#b66dff"
  )) +
  geom_jitter(
    aes(x = ssurgo_om_mean, y = Yield_decomp_mult, color = ssurgo_order_mode),
    alpha = 0.4,
    size = 1
  ) +
  # geom_smooth(
  #   data = test2,
  #   aes(x = ssurgo_om_mean, y = predicted.yield),
  #   method = "lm",
  #   color = "black"
  # ) +
  facet_grid(. ~ spei.cut_f) +
  labs(x = "SOM (%)", y = "Proportion expected yield") +
  theme(legend.title = element_blank(), legend.position = "bottom") +
  theme(
    axis.title = element_text(size = 12),
    axis.text.y = element_text(size = 9),
    axis.text.x = element_text(
      size = 9,
      angle = 45,
      hjust = 1
    ),
    legend.text = element_text(size = 10)
  ) + 
  guides(colour = guide_legend(override.aes = list(size=5, alpha = 0.8))) + 
  geom_hline(yintercept = 1, color = "grey")

dat_text <- data.frame(
  spei.cut = c("Normal","Moderate","Severe","Very severe"),
  slope = c(0.73, 0.76, 1.18, 2.2)
)

dat_text$spei.cut_f = factor(dat_text$spei.cut, levels=c("Normal","Moderate","Severe","Very severe"))

# this creates Fig. 1
yield.plot <- yield.plot + geom_text(
  size = 3.25,
  data    = dat_text,
  mapping = aes(x = 9.5, y = 14.25, label = slope),
  hjust   = "right",
  vjust   = "top"
)

```

```{r, coeff figures}


coeff.figure <- coeff.table %>%
    ungroup(.) %>%
    mutate(spei.cut= factor(spei.cut,levels=c("Normal","Moderate","Severe","Very severe"))) %>%
  filter(term == "ssurgo_om_mean") %>%
  ggplot(., aes(x = spei.cut, y = estimate)) +
  geom_errorbar(aes(ymin = estimate - std.error, ymax = estimate + std.error), width = 0) +
  geom_point()+
  xlab("Drought Level") +
  ylab("")+ 
  theme(axis.title = element_text(size=12), axis.text.y = element_text(size=10),
          axis.text.x = element_text(size=10), legend.text = element_text(size = 10))

ggsave(coeff.figure, filename = "yieldcoeff.jpg", width = 2, height = 3, units = "in")

```

```{r}

# install.packages("ggpubr")

ggpubr::ggarrange(yield.plot, coeff.figure, labels = c("A","B"), nrow = 2)

```

```{r}

coeff.table.clay <- tribble(
  ~Drought, ~Coeff, ~std.error, ~p.value,
  "Normal", summary(normal)$coefficients[3,1], summary(normal)$coefficients[3,2], summary(normal)$coefficients[3,5],
  "Moderate",summary(moderate)$coefficients[3,1], summary(moderate)$coefficients[3,2], summary(moderate)$coefficients[3,5],
  "Severe",summary(severe)$coefficients[3,1], summary(severe)$coefficients[3,2], summary(severe)$coefficients[3,5],
  "Very severe",summary(very_svr)$coefficients[3,1], summary(very_svr)$coefficients[3,2], summary(very_svr)$coefficients[3,5],
) %>%
  mutate(Drought=factor(Drought,levels=c("Normal","Moderate","Severe","Very severe")))


```

```{r}

coeff.table.clay.som <- tribble(
  ~Drought, ~Coeff, ~std.error, ~p.value,
  "Normal", summary(normal)$coefficients[4,1], summary(normal)$coefficients[4,2], summary(normal)$coefficients[4,5],
  "Moderate",summary(moderate)$coefficients[4,1], summary(moderate)$coefficients[4,2], summary(moderate)$coefficients[4,5],
  "Severe",summary(severe)$coefficients[4,1], summary(severe)$coefficients[4,2], summary(severe)$coefficients[4,5],
  "Very severe",summary(very_svr)$coefficients[4,1], summary(very_svr)$coefficients[4,2], summary(very_svr)$coefficients[4,5],
) %>%
  mutate(Drought=factor(Drought,levels=c("Normal","Moderate","Severe","Very severe")))

```

```{r}
# coeff

coeff.table <- tribble(
  ~Drought, ~Coeff, ~std.error, ~p.value,
  "Normal", summary(normal)$coefficients[2,1], summary(normal)$coefficients[2,2], summary(normal)$coefficients[2,5],
  "Moderate",summary(moderate)$coefficients[2,1], summary(moderate)$coefficients[2,2], summary(moderate)$coefficients[2,5],
  "Severe",summary(severe)$coefficients[2,1], summary(severe)$coefficients[2,2], summary(severe)$coefficients[2,5],
  "Very severe",summary(very_svr)$coefficients[2,1], summary(very_svr)$coefficients[2,2], summary(very_svr)$coefficients[2,5],
) %>%
  mutate(Drought=factor(Drought,levels=c("Normal","Moderate","Severe","Very severe")))


```

