---
title: "Yield data analysis"
output: html
---

Setup chunk: Load data and libraries. 

```{r, Setup, echo = FALSE, message = FALSE}
## libraries #####

library(tidyverse)
library(lme4)
library(mgcv)
library(itsadug)
library(reghelper)
library(wesanderson)

# data #####
all_data <- readRDS("../../data/all_data_2020.08.07.rds")
all.data.stan <- readRDS("../../data/all_data_stan_2020.08.07.rds")

```

To start we'll do a series of simple exploratory plots to see whether or not we choose to use soil data summarized using mean or geometric mean, and whether or not we choose to use SOC or SOM as our metric of choice. 

```{r, Plot yield by SOC}

ggplot(data = all_data, aes(x = ssurgo_clay_mean, y = ssurgo_h_square))+
  geom_jitter()+
  geom_smooth()

ggplot(data = all_data, aes(x = ssurgo_h, y = Yield_decomp_add))+
  geom_jitter()+
  geom_smooth()

ggplot(data = all_data, aes(x = ssurgo_om_mean, y = Yield_decomp_add))+
  geom_jitter()+
  geom_smooth()

ggplot(data = all_data, aes(x = ssurgo_clay_mean, y = Yield_decomp_add))+
  geom_jitter()+
  geom_smooth()
```
Next, we'll develop simple linear mixed models to analyze the data. 

```{r}
full.main <- all.data.stan %>%
  lmerTest::lmer(
    data = .,
    formula = Yield_decomp_add ~ summer_spei + ssurgo_om_mean + ssurgo_clay_mean +
      ssurgo_h + ssurgo_cec_mean + (1 | GEOID)
  )

car::vif(full.main)
summary(full.main)

```
VIF for CEC is very high. Remove CEC and refit. 

```{r,}

full.main <- all.data.stan %>%
  lmerTest::lmer(data = ., 
           formula = Yield_decomp_add ~ summer_spei+ssurgo_om_mean+ssurgo_clay_mean+ssurgo_h+(1|GEOID))


car::vif(full.main)
summary(full.main)

```

VIFs are low now. Final variables will be SOM, Clay, and H+ concentration. 

```{r, Yield models with SPEI, echo = TRUE}

# SPEI as drought indicator

m.state <- all.data.stan %>%
  lmerTest::lmer(data = ., 
           formula = Yield_decomp_add ~ summer_spei*ssurgo_om_mean*ssurgo_h*ssurgo_clay_mean+(1|state_alpha))

broom.mixed::tidy(m.state) %>%
  mutate(p.value = round(p.value, digits = 5))

# Plot residuals and inspect
ggplot(data.frame(eta=predict(m.state,type="link"),pearson=residuals(m.state,type="pearson")),
       aes(x=eta,y=pearson)) +
  geom_point() +
  theme_bw()


```


Log likelihood ratio tests, anova, and conditional r-squared estimates indicate that the random effect of county improves model fit.

Below we'll look at the coefficients. 

```{r}

summary(m.state)

```
There are just a few interaction effects with SPEI. We'll evaluate them here. 

```{r, SOM:SPEI 2-way interaction}

interactions::interact_plot(
  model = m.state,
  pred = ssurgo_om_mean,
  modx = summer_spei,
  modx.values = quantile(all.data.stan$summer_spei),
  partial.residuals = T
)
```


As SPEI decreases, drought severity increases, SOM effect is stronger. 

```{r, SOM:SPEI:clay 3-way interaction}

interactions::interact_plot(
  model = m.state,
  pred = ssurgo_om_mean,
  mod2 = summer_spei,
  mod2.values = quantile(all.data.stan$summer_spei),
  modx = ssurgo_clay_mean,
  modx.values = quantile(all.data.stan$ssurgo_clay_mean),
  partial.residuals = T
)
```
Clay:SOM:SPEI. Yields increase with SOM more strongly under drought. Clay amplifies this effect. This appears to be largely because of counties that are high clay, low SOM. 

```{r, SOM:SPEI:H+ 3-way interaction}

interactions::interact_plot(
  model = m.state,
  pred = ssurgo_om_mean,
  mod2 = summer_spei,
  mod2.values = quantile(all.data.stan$summer_spei),
  modx = ssurgo_h,
  modx.values = quantile(all.data.stan$ssurgo_h)[3:5],
  partial.residuals = T
)
```
OM:SPEI:H+. Yields increase with SOM more strongly under drought. As drought severity decreases, SOM has a more strongly negative effect on yields when soils are also very high H+ concentrations. This also seems to be due to very few points that are high pH.

Interaction plots indicate that SOM interacts with SPEI to mitigate the effect of low SPEI (drought) on yields. SOM has a greater positive effect on yields under all conditions when clay content is high. Higher clay content suppresses yields under drought conditions. There are additional interaction effects with SOM and SPEI for H+ and clay. The interaction with clay simply enhances the same effect, whereas the one with H+ reverses the effect at very high H+ concentrations. 

Now we'll check if the same patterns hold when we use DSCI as the drought indicator instead of SPEI. 

```{r, Yield models with SOC and DSCI, echo = TRUE}
# DSCI as drought indicator

m.state.dsci <- all.data.stan %>%
  lmerTest::lmer(data = ., 
           formula = Yield_decomp_add ~ DSCI.mean*ssurgo_om_mean*ssurgo_h*ssurgo_clay_mean+(1|state_alpha))

# View(broom.mixed::tidy(m.state.dsci) %>%
#   mutate(p.value = round(p.value, digits = 5)))

```

They mostly agree. Directionality of coeffecients is the same, as are significance factors and size of coefficients. An additional three-way interaction with DSCI:CLAY:H+ emerges. Since that does not affect the main hypotheses, we're not concerned. Given agreement between DSCI and SPEI models, we'll elect to use SPEI going forward as it introduces fewer possible artefacts from observer bias and is more strongly tied to physcial metrics.

Next, we'll generate similar plots on the subset of observations below 0 SPEI given the interesting patterns revealed by the above figures. 
```{r,}
# All conditions no SPEI 

all.data.stan %>%
  ggplot(data = ., aes(x = ssurgo_om_mean, y = Yield_decomp_add))+ 
  geom_point()

all.conditions <- all.data.stan %>%
  lmerTest::lmer(data = ., formula = Yield_decomp_add ~ ssurgo_om_mean*ssurgo_clay_mean*ssurgo_h+(1|state_alpha))

plot(all.conditions)
summary(all.conditions)

```


```{r,}
# All drought, SPEI less than mean

all.data.stan %>%
  filter(spei.cut %in% c("Very severe", "Severe", "Moderate")) %>%
  ggplot(data = ., aes(x = ssurgo_om_mean, y = Yield_decomp_add))+ 
  geom_point()

all.drought <- all.data.stan %>%
  filter(spei.cut %in% c("Very severe", "Severe", "Moderate")) %>%
  lmerTest::lmer(data = ., formula = Yield_decomp_add ~ ssurgo_om_mean*ssurgo_clay_mean*ssurgo_h+(1|state_alpha))


plot(all.drought)
summary(all.drought)



```

```{r,}
# Very severe drought

all.data.stan %>%
  filter(spei.cut == "Very severe") %>%
  ggplot(data = ., aes(x = ssurgo_h, y = Yield_decomp_add)) + geom_point() 
  
very_svr <- all.data.stan %>%
  filter(spei.cut == "Very severe") %>% 
  lmerTest::lmer(data = ., formula = Yield_decomp_add ~ ssurgo_om_mean*ssurgo_clay_mean*ssurgo_h+(1|state_alpha))

summary(very_svr)


```

```{r,}
# severe drought
all.data.stan %>%
  filter(spei.cut == "Severe") %>%
  ggplot(data = ., aes(x = ssurgo_om_mean, y = Yield_decomp_add)) + geom_point() 
  
severe <- all.data.stan %>%
  filter(spei.cut == "Severe",
         #!(ssurgo_clay_mean > 0 & ssurgo_om_mean > 1),
         #!(ssurgo_clay_mean <= -.25 & ssurgo_om_mean > 1)
         ) %>%
  lmerTest::lmer(
    data = .,
    formula = Yield_decomp_add ~ ssurgo_om_mean * ssurgo_clay_mean * ssurgo_h +
      (1 | state_alpha)
  )

summary(severe)

1537

interactions::interact_plot(
  model = severe,
  pred = ssurgo_om_mean,
  modx = ssurgo_clay_mean,
  modx.values = c(-0.5,-0.25, 0, 0.25),
  # mod2 = ssurgo_h,
  # mod2.values = c(-0.5,-0.25, 0, 0.25,1),
  plot.points = T
)

```


```{r,}

# Moderate drought, SPEI > -0.4264771, < 0.1208768

all.data.stan %>%
  #mutate(hfct = cut(ssurgo_h, breaks = quantile(all.data.stan$ssurgo_h))) %>%
  mutate(hfct = cut(ssurgo_om_mean, breaks = c(-3, -1, -0.5,0, 0.5, 1, 3.2))) %>%
  filter(spei.cut == "Moderate") %>%
  ggplot(data = ., aes(x = ssurgo_h, y = Yield_decomp_add)) + geom_point() + geom_smooth(method = "lm") + facet_wrap("hfct")
  
moderate <- all.data.stan %>%
  filter(spei.cut == "Moderate",
         #!(ssurgo_h > 0 & ssurgo_om_mean >1)
         ) %>%
  lmerTest::lmer(data = ., formula = Yield_decomp_add ~ ssurgo_om_mean*ssurgo_clay_mean*ssurgo_h+(1|state_alpha))

summary(moderate)

3998 -3956

interactions::interact_plot(
  model = moderate,
  pred = ssurgo_om_mean,
  modx = ssurgo_h,
  modx.values = quantile(all.data.stan$ssurgo_h),
plot.points = T)

```

```{r,normal to above average precip}

all.data.stan %>%
  filter(spei.cut == "Normal") %>%
  ggplot(data = ., aes(x = ssurgo_om_mean, y = Yield_decomp_add)) + geom_point() 
  
normal <- all.data.stan %>%
  filter(spei.cut == "Normal") %>%
  lmerTest::lmer(data = ., formula = Yield_decomp_add ~ ssurgo_om_mean*ssurgo_clay_mean*ssurgo_h+(1|state_alpha))

summary(normal)

interactions::interact_plot(
  model = normal,
  pred = ssurgo_h,
  modx = ssurgo_om_mean,
  modx.values = c(-0.5, 0, 0.5),
  mod2 = ssurgo_clay_mean,
  mod2.values = c(-0.5, 0, 0.5),
plot.points = T)


```
We ran models on different subsets of data to represent different levels of drought intensity. No matter the level of drought, SOC had a positive, significant effect on yields. At more intense levels of drought, that effect was more stronger, i.e. the coefficient was greater. Based on these data, we conclude that counties with higher SOC content per estimations in the gSSURGO database have greater resilence to yield losses in drought years.  


```{r}

mod <- function(df){
  lmerTest::lmer(data = df, Yield_decomp_add ~ ssurgo_om_mean*ssurgo_clay_mean*ssurgo_h+(1|state_alpha))
}

model_by_drought <- all.data.stan %>%
  group_by(spei.cut) %>%
  nest() %>%
  mutate(model_by_drought = map(data, mod))

coeff.table <- model_by_drought %>% 
  mutate(glance = map(model_by_drought, broom.mixed::tidy)) %>% 
  unnest(glance) %>%
  mutate(p.value = round(p.value, digits = 5)) %>%
  filter(effect == "fixed", 
         term != "(Intercept)") %>%
  arrange(spei.cut) %>%
  select(-data, -model_by_drought, -effect, - group)
```



```{r, Presentation figures}


test <-
  plyr::ldply(unique(all.data.stan$state_alpha), function(x) {
    data.frame(
      state_alpha = x,
      ssurgo_om_mean = seq(-3, 4, 0.05),
      ssurgo_clay_mean = mean(all.data.stan$ssurgo_clay_mean),
      ssurgo_h = mean(all.data.stan$ssurgo_h)
    )})
   
som.rev <- function(x){
  (x*(2*sd(all_data$ssurgo_om_mean)))+mean(all_data$ssurgo_om_mean)
}

test2 <- test %>%
  mutate(Normal = predict(object = normal, newdata = .), 
         Moderate = predict(object = moderate, newdata = .),
         Severe = predict(object = severe, newdata = .),
         "Very severe" = predict(object = very_svr, newdata = .),
         ssurgo_om_mean = som.rev(ssurgo_om_mean)) %>%
  filter(ssurgo_om_mean > 0, ssurgo_om_mean < 10) %>%
  select(ssurgo_om_mean, Normal, Moderate, Severe, 'Very severe') %>%
  pivot_longer(-ssurgo_om_mean, names_to = "spei.cut", values_to = "predicted.yield")

ggplot(test2, aes(x=om_descale, y=pred))+
  geom_smooth(method = "lm")




yield.plot <- all_data %>%
  mutate(spei.cut=factor(spei.cut,levels=c("Normal","Moderate","Severe","Very severe"))) %>%
  arrange(spei.cut) %>%
  ggplot(data = .) + 
  scale_color_manual(values = c("#49A942","#A0C04D",
                                "#23487A","#0096D6",
                                "#90214A","#C5351C",
                                "#F3901D"))+
  geom_jitter(aes(x = ssurgo_om_mean, y = Yield_decomp_add, color = ssurgo_order_mode), alpha = 0.2, size = 1)+
  geom_smooth(data = test2, aes(x = ssurgo_om_mean, y = predicted.yield), method = "lm", color = "black")+
  facet_wrap(facets = "spei.cut", ncol = 4)+
  labs(x = "SOM (%)", y = expression(Yield~Mg~ha^-1))+
  theme(legend.title = element_blank(), legend.position = "bottom")+
  theme(axis.title = element_text(size=12), axis.text.y = element_text(size=10),
          axis.text.x = element_text(size=10), legend.text = element_text(size = 10))

ggsave(yield.plot, filename = "~/yield.jpg", width = 8, height = 4, units = "in")


```

```{r, coeff figures}


coeff.figure <- coeff.table %>%
    ungroup(.) %>%
    mutate(spei.cut= factor(spei.cut,levels=c("Normal","Moderate","Severe","Very severe"))) %>%
  filter(term == "ssurgo_om_mean") %>%
  ggplot(., aes(x = spei.cut, y = estimate)) +
  geom_errorbar(aes(ymin = estimate - std.error, ymax = estimate + std.error), width = 0) +
  geom_point()+
  xlab("Drought Level") +
  ylab("")+ 
  theme(axis.title = element_text(size=12), axis.text.y = element_text(size=10),
          axis.text.x = element_text(size=10), legend.text = element_text(size = 10))

ggsave(coeff.figure, filename = "yieldcoeff.jpg", width = 2, height = 3, units = "in")

```

```{r}

install.packages("ggpubr")

ggpubr::ggarrange(yield.plot, coeff.figure, labels = c("A","B"), nrow = 2)

```

```{r}

coeff.table.clay <- tribble(
  ~Drought, ~Coeff, ~std.error, ~p.value,
  "Normal", summary(normal)$coefficients[3,1], summary(normal)$coefficients[3,2], summary(normal)$coefficients[3,5],
  "Moderate",summary(moderate)$coefficients[3,1], summary(moderate)$coefficients[3,2], summary(moderate)$coefficients[3,5],
  "Severe",summary(severe)$coefficients[3,1], summary(severe)$coefficients[3,2], summary(severe)$coefficients[3,5],
  "Very severe",summary(very_svr)$coefficients[3,1], summary(very_svr)$coefficients[3,2], summary(very_svr)$coefficients[3,5],
) %>%
  mutate(Drought=factor(Drought,levels=c("Normal","Moderate","Severe","Very severe")))


```
```{r}

coeff.table.clay.som <- tribble(
  ~Drought, ~Coeff, ~std.error, ~p.value,
  "Normal", summary(normal)$coefficients[4,1], summary(normal)$coefficients[4,2], summary(normal)$coefficients[4,5],
  "Moderate",summary(moderate)$coefficients[4,1], summary(moderate)$coefficients[4,2], summary(moderate)$coefficients[4,5],
  "Severe",summary(severe)$coefficients[4,1], summary(severe)$coefficients[4,2], summary(severe)$coefficients[4,5],
  "Very severe",summary(very_svr)$coefficients[4,1], summary(very_svr)$coefficients[4,2], summary(very_svr)$coefficients[4,5],
) %>%
  mutate(Drought=factor(Drought,levels=c("Normal","Moderate","Severe","Very severe")))

```

```{r}
# coeff

coeff.table <- tribble(
  ~Drought, ~Coeff, ~std.error, ~p.value,
  "Normal", summary(normal)$coefficients[2,1], summary(normal)$coefficients[2,2], summary(normal)$coefficients[2,5],
  "Moderate",summary(moderate)$coefficients[2,1], summary(moderate)$coefficients[2,2], summary(moderate)$coefficients[2,5],
  "Severe",summary(severe)$coefficients[2,1], summary(severe)$coefficients[2,2], summary(severe)$coefficients[2,5],
  "Very severe",summary(very_svr)$coefficients[2,1], summary(very_svr)$coefficients[2,2], summary(very_svr)$coefficients[2,5],
) %>%
  mutate(Drought=factor(Drought,levels=c("Normal","Moderate","Severe","Very severe")))


```

