---
title: "nass-data-processing-ML_test"
output: html_document
date: "2023-03-15"
editor_options: 
  chunk_output_type: console
---

# setup
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(rnassqs)
library(dplyr)
library(ggplot2)
library(tidyr)
library(reshape2)
library(caret)
library(parallel)
library(readr)
api_key <- as.character(read.csv("code/NASS_API_key.csv", header = F)[1,1])       # api key

# Call in all corn yield data via NASS API ####

nassqs_auth(key = api_key)

```

# Pull data from NASS   ##### OR YOU CAN SKIP THIS AND GO TO LINE 200 "d.1 <- read_rds(...)" ##### as of 3/29/23

1. Specify years and pull yield data from nass 'SURVEY'

```{r yield}
# Specify the range of years across which you want to collect data
years <- as.list(2000:2022)  
# ML: changed to 2000:2022 to include more data points
# 2000-2022: 38847 obs
# 2000-2016: 30108 obs

# BM: Future self, when update with newer data: consider which years to use to minimize 
# the confounding effect of improvements in maize genetics, i.e., shift the
# 16 year window forward, drop oldest records.

## Yields

d <- plyr::ldply(years, function(x){
  
  params <- list(
    commodity_desc = "CORN",
    util_practice_desc = "GRAIN",
    year = x,
    agg_level_desc = "COUNTY", 
    source_desc = "SURVEY",  # change source to source_desc
    domain_desc = "TOTAL"
  )
  
  return(
    nassqs_yields(params) %>%
      filter(
        prodn_practice_desc == "ALL PRODUCTION PRACTICES",
        county_ansi != ""
      ) %>%
      mutate(
        GEOID = paste(state_ansi, county_ansi, sep = ""),
        Yield_mg_ha = as.numeric(Value) * 0.0628
      ) %>%
      dplyr::select(
        year,
        GEOID,
        state_alpha,
        state_ansi,
        county_ansi,
        county_name,
        Yield_mg_ha
      )
  )
})

#ATR addition: write the pulled nass data as it's own data file so we no longer have to pull it every time
#write_rds(d, path = "data/nass_03142023.rds")
```

2. Calculate irrigation acres from census data and create a filter
Note: we can skip this step if we'd like to keep data from irrigated land (Aysha)
Meng: added the 2017 new census year

```{r irrgation}
### Total acres
census.years <- as.list(c(1997,2002,2007,2012, 2017))

d.acres.total <- plyr::ldply(census.years, function(x) {
  
  params <- list(
    commodity_desc = "CORN",
    util_practice_desc = "GRAIN",
    source_desc = "CENSUS",
    year = x,
    agg_level_desc = "COUNTY",
    short_desc = "CORN, GRAIN - ACRES HARVESTED",
    domain_desc = "TOTAL"
  )
  
  return(
    nassqs(params) %>%
      filter(county_ansi != "") %>%
      mutate(
        GEOID = paste(state_ansi, county_ansi, sep = ""),
        Acres_total = as.numeric(gsub(
          Value, pattern = ",", replacement = ""
        ))
      ) %>%
      select(
        year,
        GEOID,
        state_alpha,
        state_ansi,
        county_ansi,
        county_name,
        Acres_total
      )
  )
})

##### IRRIGATED ACRES

d.acres.irrigated <- plyr::ldply(census.years, function(x) {
  
  params <- list(
    commodity_desc = "CORN",
    util_practice_desc = "GRAIN",
    source_desc = "CENSUS",
    year = x,
    agg_level_desc = "COUNTY",
    short_desc = "CORN, GRAIN, IRRIGATED - ACRES HARVESTED",
    domain_desc = "TOTAL"
  )
  
  return(
    nassqs(params) %>%
      filter(county_ansi != "") %>%
      mutate(
        GEOID = paste(state_ansi, county_ansi, sep = ""),
        Acres_irrigated = as.numeric(gsub(
          Value, pattern = ",", replacement = ""
        ))
      ) %>%
      select(
        year,
        GEOID,
        state_alpha,
        state_ansi,
        county_ansi,
        county_name,
        Acres_irrigated
      )
  )
})

##

d.acres <- d.acres.total %>%
  left_join(d.acres.irrigated) %>%
  filter(GEOID %in% d$GEOID,!is.na(Acres_total)) %>%
  replace_na(list(Acres_irrigated = 0)) %>%
  mutate(Percent_irrigated = Acres_irrigated / Acres_total) %>%
  group_by(GEOID) %>%
  summarize(
    Mean.percent.irrigated = mean(Percent_irrigated),
    SD.percent.irrigated = sd(Percent_irrigated)
  )

## FILTER BASED ON IRRIGATED ACRES DATA

# Create filter to select counties that are 5 percent or less irrigated, 
# choice of 5 percent based on dsitribution of percentages, vast majority of counties are 5 percent or less irrigated

d.irrgiated.filter <- d.acres %>%
  filter(Mean.percent.irrigated <= 0.05) %>%
  filter(SD.percent.irrigated <= 0.01) 


```

3. filter yield data based on the availability of year numbers
ML: Instead of removing data from irrigated land, a new variable 'irrigated' was created:
irrigated == 'irrigated' means more than 95% of acres of a county were irrigated
irrigated == 'rainfed' means less than 5% of acres of a county were irrigated

This step is just for testing different filters. Once the filters are chosen, this step can be skipped. 

```{r filter}
# ML: check how many data points are removed using different number of years
d.1 <- d %>%
  mutate(irrigated = case_when(GEOID %in% d.irrgiated.filter$GEOID ~ 'rainfed',
                               TRUE ~ 'irrigated')) %>%
  add_count(irrigated) %>%  # irrigated 18111; rainfed 20736
  group_by(state_alpha,GEOID,irrigated) %>%
  summarise(total_n =n()) 

# save d.1
# write_rds(d.1, file="data/d1.rds")
d.1 <- read_rds("data/d1.rds")
# open a graphics window
windows(xpinch=200, ypinch=200, width=5, height=5)

# re-level so that not irrigated is on the bottom, irrigated on the top
d.1$irrigated <- factor(d.1$irrigated)
levels(d.1$irrigated)
d.1$irrigated <- relevel(d.1$irrigated, "Y")
levels(d.1$irrigated)

# plot all data together
# need this to be a named object, p1, to make cumulative plot below
p1 <- ggplot(d.1, aes(x = total_n,fill= irrigated)) +
  geom_histogram( binwidth = 1,col="grey",position = "stack") +
  scale_x_continuous(breaks = seq(1,23,1)) +
  scale_y_continuous(breaks = c(0,50,100,200,300,400,500)) +
  labs(x = 'Number of years', y = 'Number of counties') 

p1 

### make histogram-like chart showing how filter by year affects amount of data
# need to show that if we set the filter to minimum 1 year, you have all the counties with 
# 1 to 23 years of corn yield data s(biggest bar)
# if minimum set to 23 years all you have is counties with 23 years (smallest bar)

p1.d <- ggplot_build(p1)$data[[1]][, c(3:4,12)] %>%   # extract data from p1 above
# $x is filter for these many years
# $group is irrigated, 1=no, 2=yes
# we want to make new column that, within groups, sums the years before it
        group_by(group) %>%   # group by rainfed/irrigated
        map_df(rev) %>%       # reverse the order of the rows so that the cumulative sums in next step do what we want
        mutate(cum.count=cumsum(count)) %>%  # cumulative sums
        map_df(rev)   # put the x's back in the right order for plotting

p1.d$group <- relevel(factor(p1.d$group), "2")

# cumulative plot (all states together, no facets)
ggplot(p1.d,aes(x = x, 
                y = cum.count, 
                fill = group)) +
  geom_bar(stat="identity", position = "stack") +
  scale_x_continuous(breaks = seq(1,23,1), expand=c(0,0)) +
  scale_y_continuous(expand=c(0,0)) +
  scale_fill_manual(values=c("#40b0a6", "#e1be6a"),  # using a colorblind friendly color scheme
                    breaks=c("2", "1"),
                    labels=c("Irrigated", "Rainfed"),
                    name="Key") +
  labs(x = 'Filter: County minimum number of corn years', y = 'Number of county-years') +
  theme(panel.grid.minor=element_blank(), 
        panel.grid.major=element_blank() ,
        panel.background = element_rect(fill = NA) ,
        panel.border=element_rect(color="grey50", fill=NA, linewidth=0.5),
        strip.text=element_text(size=rel(1), face="bold", vjust=-2),
        strip.background = element_rect(fill=NA),
        panel.margin=unit(0, "lines"),
        panel.spacing.y=unit(-1, "points"),
        legend.text=element_text(size=14),
        legend.title=element_text(size=14),
        axis.title.x = element_text(margin=margin(8, 0,0,0)),
        axis.title.y = element_text(margin=margin(0, 8,0,0))                        
        # legend.title=element_text(size=12, face="bold"),
        # plot.margin = unit(c(0.5,0.5,0.5,0.5), "cm")
  )	

ggsave("code/plots/filters_cumulative_years.png")


# plot data by states - just rainfed
p2 <- ggplot(d.1,aes(x = total_n,fill= irrigated)) +  # naming object to extract data for cumulative plot below
  geom_histogram(position = "stack",binwidth = 1) +
  scale_x_continuous(breaks = seq(1,23,5)) +
  scale_fill_manual(values=c("#00000000", "#e1be6a"),   # make irrigated invisible for now, using a colorblind friendly color scheme 
                    breaks=c("Y", "N"),
                    labels=c("Irrigated", "Rainfed"),
                    name="Key") +
  labs(x = 'Number of years', y = 'Number of counties') +
  facet_wrap(~state_alpha) +
  theme(panel.grid.minor=element_blank(), 
        panel.grid.major=element_blank() ,
        panel.background = element_rect(fill = NA) ,
        panel.border=element_rect(color="grey50", fill=NA, linewidth=0.5),
        strip.text=element_text(size=rel(1), face="bold", vjust=-2),
        strip.background = element_rect(fill=NA),
        panel.spacing.x=unit(0, "lines"),
        panel.spacing.y=unit(-1, "points"),
        legend.text=element_text(size=14),
        legend.title=element_text(size=14)
        # legend.title=element_text(size=12, face="bold"),
        # plot.margin = unit(c(0.5,0.5,0.5,0.5), "cm")
        
  )	

p2

ggsave("code/plots/filters_years_rainfed_only.png")

# plot data by states - rainfed and irrigated
ggplot(d.1,aes(x = total_n,fill= irrigated)) +
  geom_histogram(position = "stack",binwidth = 1) +
  scale_x_continuous(breaks = seq(1,23,5)) +
  scale_fill_manual(values=c("#40b0a6", "#e1be6a"),  # using a colorblind friendly color scheme 
                    breaks=c("Y", "N"),
                    labels=c("Irrigated", "Rainfed"),
                    name="Key") +
  labs(x = 'Number of years', y = 'Number of counties') +
  facet_wrap(~state_alpha) +
  theme(panel.grid.minor=element_blank(), 
        panel.grid.major=element_blank() ,
        panel.background = element_rect(fill = NA) ,
        panel.border=element_rect(color="grey50", fill=NA, linewidth=0.5),
        strip.text=element_text(size=rel(1), face="bold", vjust=-2),
        strip.background = element_rect(fill=NA),
        panel.margin=unit(0, "lines"),
        panel.spacing.y=unit(-1, "points"),
        legend.text=element_text(size=14),
        legend.title=element_text(size=14)
        # legend.title=element_text(size=12, face="bold"),
        # plot.margin = unit(c(0.5,0.5,0.5,0.5), "cm")
        
  )	 

ggsave("code/plots/filters_years_rainfed_and_irr.png")


### arrange the data to make cumulative plots by state --
### this creates the text used in the mutate() step below, only needed once
# states <- unique(d.1$state_alpha)
# panel <- seq(1,41,1)
# states_key <- data.frame(states=states, panel=panel)
# states_key$panel <- as.character(states_key$panel)
# # make the string we'll use in relevel. Note \" within " " tells R that you 
# # want to show the double quotes in the string
# states_key$key <- paste0("\"",states_key$panel, "\"=\"", states_key$states, "\"")
# # check our work
# key <- stringr::str_flatten_comma(states_key$key)
# cat(key)

p2.d <- ggplot_build(p2)$data[[1]][, c(3:4,11,12)] %>%   # extract data from p1 above
# $x is filter for these many years
# $group is irrigated, 1=yes, 2=no ---why is it opposite here than in p1.d??? IDK
# $panel is state
# we want to make new column that, within groups, sums the years before it
        mutate(state=recode_factor(PANEL, "1"="AL", "2"="AR", "3"="AZ", "4"="CA", "5"="CO", "6"="DE", "7"="FL", "8"="GA", "9"="IA", "10"="ID", "11"="IL", "12"="IN", "13"="KS", "14"="KY", "15"="LA", "16"="MD", "17"="MI", "18"="MN", "19"="MO", "20"="MS", "21"="MT", "22"="NC", "23"="ND", "24"="NE", "25"="NJ", "26"="NM", "27"="NY", "28"="OH", "29"="OK", "30"="OR", "31"="PA", "32"="SC", "33"="SD", "34"="TN", "35"="TX", "36"="UT", "37"="VA", "38"="WA", "39"="WI", "40"="WV", "41"="WY")) %>%
        mutate(irr=recode_factor(group, "1"="Irrigated", "2"="Rainfed")) %>%
        group_by(state,irr) %>% # group by rainfed/irrigated
        arrange(desc(x), .by_group=T) %>%       # reverse the order of the rows so that the cumulative sums in next step do what we want
        mutate(cum.count=cumsum(count)) %>%  # cumulative sums
        arrange(x, .by_group=T)   # put the x's back in the right order for plotting

p2.d$irr <- relevel(factor(p2.d$irr), "Irrigated")



# plot cumulative data by states - just rainfed
ggplot(p2.d,aes(x = x, 
                y = cum.count, 
                fill = irr)) +
  geom_bar(stat="identity", position = "stack", width=0.2) +
  scale_x_continuous(breaks = seq(0,23,5), expand=c(0,0)) +
  scale_y_continuous(expand=c(0,0)) +
  scale_fill_manual(values=c("#00000000", "#e1be6a"),  # make irrigated invisible for now, using a colorblind friendly color scheme 
                    name="Key") +
  facet_wrap(~state) +
  labs(x = 'Filter: County minimum number of corn years', y = 'Number of county-years') +
  theme(panel.grid.minor=element_blank(), 
        panel.grid.major=element_blank() ,
        panel.background = element_rect(fill = NA) ,
        panel.border=element_rect(color="grey50", fill=NA, linewidth=0.5),
        strip.text=element_text(size=rel(1), face="bold", vjust=-2),
        strip.background = element_rect(fill=NA),
        panel.margin=unit(0, "lines"),
        panel.spacing.y=unit(-1, "points"),
        legend.text=element_text(size=14),
        legend.title=element_text(size=14),
        axis.title.x = element_text(margin=margin(8, 0,0,0)),
        axis.title.y = element_text(margin=margin(0, 8,0,0))                        
        # legend.title=element_text(size=12, face="bold"),
        # plot.margin = unit(c(0.5,0.5,0.5,0.5), "cm")
  )	

ggsave("code/plots/filters_years_cumulative_rainfed_only.png")


# plot cumulative data by states - rainfed and irrigated
ggplot(p2.d,aes(x = x, 
                y = cum.count, 
                fill = irr)) +
  geom_bar(stat="identity", position = "stack", width=0.2) +
  scale_x_continuous(breaks = seq(0,23,5), expand=c(0,0)) +
  scale_y_continuous(expand=c(0,0)) +
  scale_fill_manual(values=c("#40b0a6", "#e1be6a"), 
                    name="Key") +
  facet_wrap(~state) +
  labs(x = 'Filter: County minimum number of corn years', y = 'Number of county-years') +
  theme(panel.grid.minor=element_blank(), 
        panel.grid.major=element_blank() ,
        panel.background = element_rect(fill = NA) ,
        panel.border=element_rect(color="grey50", fill=NA, linewidth=0.5),
        strip.text=element_text(size=rel(1), face="bold", vjust=-2),
        strip.background = element_rect(fill=NA),
        panel.margin=unit(0, "lines"),
        panel.spacing.y=unit(-1, "points"),
        legend.text=element_text(size=14),
        legend.title=element_text(size=14),
        axis.title.x = element_text(margin=margin(8, 0,0,0)),
        axis.title.y = element_text(margin=margin(0, 8,0,0))                        
        # legend.title=element_text(size=12, face="bold"),
        # plot.margin = unit(c(0.5,0.5,0.5,0.5), "cm")
  )	

ggsave("code/plots/filters_years_cumulative_rainfed_and_irr.png")

```


4. apply filters

irrigated == 'irrigated' means more than 95% of acres of a county were irrigated
irrigated == 'rainfed' means less than 5% of acres of a county were irrigated

Tried two filters for number of years: 

n = 15 (from Dan's code)
n = 12 (half of years 2000-2022)

```{r apply_filters}
d.new.15 <- d %>%
  mutate(irrigated = case_when(GEOID %in% d.irrgiated.filter$GEOID ~ 'rainfed',
                               TRUE ~ 'irrigated')) %>%
  group_by(GEOID) %>%
  add_count(GEOID) %>%
  filter(n >= 15) %>% # Filter to >=15 corn yield observations #ATR: This is where he resricts the acres to 15 years of corn
  ungroup(.) %>%
  select(-n) #34003 obs

d.new.12 <- d %>%
  mutate(irrigated = case_when(GEOID %in% d.irrgiated.filter$GEOID ~ 'rainfed',
                               TRUE ~ 'irrigated')) %>%
  group_by(GEOID) %>%
  add_count(GEOID) %>%
  filter(n >= 12) %>% # Filter to >=15 corn yield observations #ATR: This is where he resricts the acres to 15 years of corn
  ungroup(.) %>%
  select(-n) #35866 obs

```

5. extract de-trended yield

Note: used gamloess method to model the relationship between yield and year. 
The model was run for different counties

This step was done on server

```{r detrend}
# install.packages("tictoc")
library(tictoc) # this is to record the running time

mod <- function(df){
  df <- df
  
  grid <- expand.grid(span = seq(0.3, 0.5, len = 5), degree = seq(0,1, len=2) )
  
  grid.control <- trainControl(
    method = "repeatedcv",
    number = 10,
    repeats = 5,
    search = "grid")

  train_loess <- caret::train(Yield_mg_ha ~ year, 
                       method = "gamLoess",
                       tuneGrid=grid,
                       trControl=grid.control,
                       data = df)
  
  df$Detrend_resids <- as.numeric(residuals(train_loess))
  df$Detrend_predictions <- as.numeric(predict(train_loess))
  return(df)
}

#---------------n = 15---------------#
d.new.15$year <- as.integer(d.new.15$year) # year needs to be converted to integer to run the regression model in the train function above

# Dan's code using mclapply function
d_list_15 <- split(d.new.15, f = d.new.15$GEOID) # ML: it creates a list based on county

tic()
d_list_15.1 <- mclapply(X = d_list_15,FUN = mod, mc.cores = 20) 
toc()
# 375.489 sec elapsed

d.15 <- dplyr::bind_rows(d_list_15.1)

d.15.1 <- d.15 %>%
  group_by(GEOID) %>%
  mutate(County_avg_yield = mean(Yield_mg_ha)) %>%
  ungroup(.) %>%
  mutate(Yield_decomp_add = County_avg_yield+Detrend_resids,  # de-trended yield 
         Yield_decomp_mult = Yield_mg_ha/Detrend_predictions) # yield anomaly

# saved the data to a rds file to be used for downstream processing
write_rds(d.15.1, file = "data/corn_yield_2000-2022_w_irrigation_n_15.rds")

#---------------n = 12---------------#
d.new.12$year <- as.integer(d.new.12$year) # year needs to be converted to integer to run the regression model in the train function above

# Dan's code using mclapply function
d_list_12 <- split(d.new.12, f = d.new.12$GEOID) # ML: it creates a list based on county

tic()
d_list_12.1 <- mclapply(X = d_list_12,FUN = mod, mc.cores = 28) 
toc()

d.12 <- dplyr::bind_rows(d_list_12.1)

d.12.1 <- d.12 %>%
  group_by(GEOID) %>%
  mutate(County_avg_yield = mean(Yield_mg_ha)) %>%
  ungroup(.) %>%
  mutate(Yield_decomp_add = County_avg_yield+Detrend_resids,  # de-trended yield 
         Yield_decomp_mult = Yield_mg_ha/Detrend_predictions) # yield anomaly

# saved the data to a rds file to be used for downstream processing
write_rds(d.12.1, file = "data/corn_yield_2000-2022_w_irrigation_n_12.rds")
```


