---
title: "ATR nass data modifications"
output: html_document
date: "2023-03-23"
---


ATR: I moved Dan's nass-data-processing code into a new file to include irrigated data. I am not manipulating years at this point 
```{r}
library(rnassqs)
library(dplyr)
library(ggplot2)
library(tidyr)
library(reshape2)
library(caret)
library(parallel)
library(readr)
library(tictoc)
```

ATR: I created a data file of the pulled corn yield nass data from 2022, so I did not include the pulling of data using the API key. 

ATR revised: I added the nass api key code back in. He utilizes the nass API key again in later code. We do not have to run the Yield pulling code again though


api_key <- as.character(read.csv("C:/Users/AyshaTappRoss/Github/soilc-midwest-DanKane/code/NASS_API_key.csv", header = F)[1,1])       # api key
# Specify the range of years across which you want to collect data
years <- as.list(2000:2016)  # BM: Future self, when update with newer data: consider which years to use to minimize 
# the confounding effect of improvements in maize genetics, i.e., shift the
# 16 year window forward, drop oldest records.

# Call in all corn yield data via NASS API ####

nassqs_auth(key = api_key)


ATR: instead of using the nass data for 16 years from Dan's code I used the 22yr timeline from Meng's code 
```{r}
d <-read_rds("data/nass_22yr_03202023.rds")
```

ATR: At this point d has 38847 observations

### Converting the data to Total acres. He pulls from the nass data using the api key again in this code. I added the key back into the code at the top. I don't think we need this step though. He uses this to create a percentage of irrigated acres of the total acres.

```{r}
api_key <- as.character(read.csv("code/NASS_API_key.csv", header = F)[1,1])       # api key
# Specify the range of years across which you want to collect data
years <- as.list(2000:2022)  #We chose to extend the years to the latest year that would be available. this is the same as the RDS file being called above

# Call in all corn yield data via NASS API ####

nassqs_auth(key = api_key)

```
##The next block is from Meng's nass_data_processing_ML_test file

2. Calculate irrigation acres from census data and create a filter
Note: we can skip this step if we'd like to keep data from irrigated land (Aysha)
Meng: I added the 2017 new census year

```{r irrgation}
### Total acres
census.years <- as.list(c(1997,2002,2007,2012, 2017))

d.acres.total <- plyr::ldply(census.years, function(x) {
  
  params <- list(
    commodity_desc = "CORN",
    util_practice_desc = "GRAIN",
    source_desc = "CENSUS",
    year = x,
    agg_level_desc = "COUNTY",
    short_desc = "CORN, GRAIN - ACRES HARVESTED",
    domain_desc = "TOTAL"
  )
  
  return(
    nassqs(params) %>%
      filter(county_ansi != "") %>%
      mutate(
        GEOID = paste(state_ansi, county_ansi, sep = ""),
        Acres_total = as.numeric(gsub(
          Value, pattern = ",", replacement = ""
        ))
      ) %>%
      dplyr::select(
        year,
        GEOID,
        state_alpha,
        state_ansi,
        county_ansi,
        county_name,
        Acres_total
      )
  )
})

##### IRRIGATED ACRES

d.acres.irrigated <- plyr::ldply(census.years, function(x) {
  
  params <- list(
    commodity_desc = "CORN",
    util_practice_desc = "GRAIN",
    source_desc = "CENSUS",
    year = x,
    agg_level_desc = "COUNTY",
    short_desc = "CORN, GRAIN, IRRIGATED - ACRES HARVESTED",
    domain_desc = "TOTAL"
  )
  
  return(
    nassqs(params) %>%
      filter(county_ansi != "") %>%
      mutate(
        GEOID = paste(state_ansi, county_ansi, sep = ""),
        Acres_irrigated = as.numeric(gsub(
          Value, pattern = ",", replacement = ""
        ))
      ) %>%
      dplyr::select(
        year,
        GEOID,
        state_alpha,
        state_ansi,
        county_ansi,
        county_name,
        Acres_irrigated
      )
  )
})

##

d.acres <- d.acres.total %>%
  left_join(d.acres.irrigated) %>%
  filter(GEOID %in% d$GEOID,!is.na(Acres_total)) %>%
  replace_na(list(Acres_irrigated = 0)) %>%
  mutate(Percent_irrigated = Acres_irrigated / Acres_total) %>%
  group_by(GEOID) %>%
  summarize(
    Mean.percent.irrigated = mean(Percent_irrigated),
    SD.percent.irrigated = sd(Percent_irrigated)
  )

## FILTER BASED ON IRRIGATED ACRES DATA

# Create filter to select counties that are 5 percent or less irrigated, 
# choice of 5 percent based on dsitribution of percentages, vast majority of counties are 5 percent or less irrigated

d.irrgiated.filter <- d.acres %>%
  filter(Mean.percent.irrigated <= 0.05) %>%
  filter(SD.percent.irrigated <= 0.01) 


```

3. filter yield data based on the availability of year numbers
ML: Instead of removing data from irrigated land, a new variable 'irrigated' was created:
irrigated == Y means more than 95% of acres of a county were irrigated
irrigated == N means less than 5% of acres of a county were irrigated

```{r filter}
# ML: check how many data points are removed using different number of years
d.1 <- d %>%
  mutate(irrigated = case_when(GEOID %in% d.irrgiated.filter$GEOID ~ 'N',
                               TRUE ~ 'Y')) %>%
  add_count(irrigated) %>%  # irrigated 18111; non-irrigated 20736
  group_by(state_alpha,GEOID,irrigated) %>%
  summarise(total_n =n()) 

# plot all data together
ggplot(d.1, aes(x = total_n,fill= irrigated)) +
  geom_histogram( binwidth = 1,col="grey",position = "stack") +
  scale_x_continuous(breaks = seq(1,23,1)) +
  scale_y_continuous(breaks = c(0,50,100,200,300,400,500)) +
  labs(x = 'Number of years', y = 'Number of counties') 

# plot data by states
ggplot(d.1,aes(x = total_n,fill= irrigated)) +
  geom_histogram(position = "stack",binwidth = 1, col="grey") +
  scale_x_continuous(breaks = seq(1,23,5)) +
  labs(x = 'Number of years', y = 'Number of counties') +
  facet_wrap(~state_alpha)

# --------------apply filters--------------#
d.new <- d %>%
  mutate(irrigated = case_when(GEOID %in% d.irrgiated.filter$GEOID ~ 'N',
                               TRUE ~ 'Y')) %>%
  group_by(GEOID) %>%
  add_count(GEOID) %>%
  filter(n >= 15) %>% # Filter to >=15 corn yield observations #ATR: This is where he resricts the acres to 15 years of corn
  ungroup(.) %>%
  dplyr::select(-n)

```


ATR original: I origially started with n=13, but after Meng created the graphs we decided to run the data starting with n=9 and report out 

ATR 3/28: Now I am going to take the data and assess each filtered year by total data point and irrigated/ non-irrigated data points. We will do this by replacing Dan's filtering code with Meng's new code

  
```{r}
d9 <- d %>%
  group_by(GEOID) %>%
  add_count(GEOID) %>%
  filter(n >= 9) %>% # Filter to >=9 corn yield observations
  ungroup(.) %>%
  dplyr::select(-n)


aggregate(d9$state_ansi, by = list(d9$state_ansi), FUN = length)
aggregate(d9$GEOID, by = list(d9$GEOID), FUN = length)

#Meng's code
d.new.9 <- d %>%
  mutate(irrigated = case_when(GEOID %in% d.irrgiated.filter$GEOID ~ 'N',
                               TRUE ~ 'Y')) %>%
  group_by(GEOID) %>%
  add_count(GEOID) %>%
  filter(n >= 9) %>% # Filter to >=9 corn yield observations #ATR: This is where he resricts the acres to 15 years of corn
  ungroup(.) %>%
  dplyr::select(-n)

d.new.N9 <-subset(d.new.9, irrigated == 'N')
d.new.Y9 <-subset(d.new.9, irrigated == 'Y')
aggregate(d.new.N9$state_ansi, by = list(d.new.N9$state_ansi, d.new.N9$irrigated ), FUN = length)
aggregate(d.new.N9$GEOID, by = list(d.new.N9$GEOID, d.new.N9$irrigated), FUN = length)
aggregate(d.new.Y9$state_ansi, by = list(d.new.Y9$state_ansi, d.new.Y9$irrigated ), FUN = length)
aggregate(d.new.Y9$GEOID, by = list(d.new.Y9$GEOID, d.new.Y9$irrigated), FUN = length)

``` 
  9= 41 States and 1914 counties 
  29 states, 1012 counties = N; 40 states, 902 counties = Y
  
```{r}
d10 <- d %>%
  group_by(GEOID) %>%
  add_count(GEOID) %>%
  filter(n >= 10) %>% # Filter to >=10 corn yield observations
  ungroup(.) %>%
  dplyr::select(-n)


aggregate(d10$state_ansi, by = list(d10$state_ansi), FUN = length)
aggregate(d10$GEOID, by = list(d10$GEOID), FUN = length)

#Meng's code
d.new.10 <- d %>%
  mutate(irrigated = case_when(GEOID %in% d.irrgiated.filter$GEOID ~ 'N',
                               TRUE ~ 'Y')) %>%
  group_by(GEOID) %>%
  add_count(GEOID) %>%
  filter(n >= 10) %>% # Filter to >=10 corn yield observations #ATR: This is where he resricts the acres to 15 years of corn
  ungroup(.) %>%
  dplyr::select(-n)

d.new.N10 <-subset(d.new.10, irrigated == 'N')
d.new.Y10 <-subset(d.new.10, irrigated == 'Y')
aggregate(d.new.N10$state_ansi, by = list(d.new.N10$state_ansi, d.new.N10$irrigated ), FUN = length)
aggregate(d.new.N10$GEOID, by = list(d.new.N10$GEOID, d.new.N10$irrigated), FUN = length)
aggregate(d.new.Y10$state_ansi, by = list(d.new.Y10$state_ansi, d.new.Y10$irrigated ), FUN = length)
aggregate(d.new.Y10$GEOID, by = list(d.new.Y10$GEOID, d.new.Y10$irrigated), FUN = length)
``` 
10: 40 states, 1848 counties
29 states, 985 counties = N
39 states, 873 counties = Y

  
```{r}
d11 <- d %>%
  group_by(GEOID) %>%
  add_count(GEOID) %>%
  filter(n >= 11) %>% # Filter to >=11 corn yield observations
  ungroup(.) %>%
   dplyr::select(-n)


aggregate(d11$state_ansi, by = list(d11$state_ansi), FUN = length)
aggregate(d11$GEOID, by = list(d11$GEOID), FUN = length)

#Meng's code
d.new.11 <- d %>%
  mutate(irrigated = case_when(GEOID %in% d.irrgiated.filter$GEOID ~ 'N',
                               TRUE ~ 'Y')) %>%
  group_by(GEOID) %>%
  add_count(GEOID) %>%
  filter(n >= 11) %>% # Filter to >=11 corn yield observations #ATR: This is where he resricts the acres to 15 years of corn
  ungroup(.) %>%
  dplyr::select(-n)

d.new.N11 <-subset(d.new.11, irrigated == 'N')
d.new.Y11 <-subset(d.new.11, irrigated == 'Y')
aggregate(d.new.N11$state_ansi, by = list(d.new.N11$state_ansi, d.new.N11$irrigated ), FUN = length)
aggregate(d.new.N11$GEOID, by = list(d.new.N11$GEOID, d.new.N11$irrigated), FUN = length)
aggregate(d.new.Y11$state_ansi, by = list(d.new.Y11$state_ansi, d.new.Y11$irrigated ), FUN = length)
aggregate(d.new.Y11$GEOID, by = list(d.new.Y11$GEOID, d.new.Y11$irrigated), FUN = length)
``` 

11: 37 states, 1818 counties
28 states, 969 counties= N; 36 states, 849 counties = Y
  
```{r}
d12 <- d %>%
  group_by(GEOID) %>%
  add_count(GEOID) %>%
  filter(n >= 12) %>% # Filter to >=12 corn yield observations
  ungroup(.) %>%
  dplyr::select(-n)


aggregate(d12$state_ansi, by = list(d12$state_ansi), FUN = length)
aggregate(d12$GEOID, by = list(d12$GEOID), FUN = length)

#Meng's code
d.new.12 <- d %>%
  mutate(irrigated = case_when(GEOID %in% d.irrgiated.filter$GEOID ~ 'N',
                               TRUE ~ 'Y')) %>%
  group_by(GEOID) %>%
  add_count(GEOID) %>%
  filter(n >= 12) %>% # Filter to >=12 corn yield observations #ATR: This is where he resricts the acres to 15 years of corn
  ungroup(.) %>%
  dplyr::select(-n)

d.new.N12 <-subset(d.new.12, irrigated == 'N')
d.new.Y12 <-subset(d.new.12, irrigated == 'Y')
aggregate(d.new.N12$state_ansi, by = list(d.new.N12$state_ansi, d.new.N12$irrigated ), FUN = length)
aggregate(d.new.N12$GEOID, by = list(d.new.N12$GEOID, d.new.N12$irrigated), FUN = length)
aggregate(d.new.Y12$state_ansi, by = list(d.new.Y12$state_ansi, d.new.Y12$irrigated ), FUN = length)
aggregate(d.new.Y12$GEOID, by = list(d.new.Y12$GEOID, d.new.Y12$irrigated), FUN = length)
``` 

12: 37 states, 1782 counties
28 states, 951 counties = N
36 states, 831 counties = Y

```{r}
d13 <- d %>%
  group_by(GEOID) %>%
  add_count(GEOID) %>%
  filter(n >= 13) %>% # Filter to >=13 corn yield observations
  ungroup(.) %>%
  dplyr::select(-n)


aggregate(d13$state_ansi, by = list(d13$state_ansi), FUN = length)
aggregate(d13$GEOID, by = list(d13$GEOID), FUN = length)

#Meng's code
d.new.13 <- d %>%
  mutate(irrigated = case_when(GEOID %in% d.irrgiated.filter$GEOID ~ 'N',
                               TRUE ~ 'Y')) %>%
  group_by(GEOID) %>%
  add_count(GEOID) %>%
  filter(n >= 13) %>% # Filter to >=13 corn yield observations #ATR: This is where he resricts the acres to 13 years of corn
  ungroup(.) %>%
  dplyr::select(-n)

d.new.N13 <-subset(d.new.13, irrigated == 'N')
d.new.Y13 <-subset(d.new.13, irrigated == 'Y')
aggregate(d.new.N13$state_ansi, by = list(d.new.N13$state_ansi, d.new.N13$irrigated ), FUN = length)
aggregate(d.new.N13$GEOID, by = list(d.new.N13$GEOID, d.new.N13$irrigated), FUN = length)
aggregate(d.new.Y13$state_ansi, by = list(d.new.Y13$state_ansi, d.new.Y13$irrigated ), FUN = length)
aggregate(d.new.Y13$GEOID, by = list(d.new.Y13$GEOID, d.new.Y13$irrigated), FUN = length)
``` 

ATR: This may be too liberal of a filter, but it gives us 35410 observations with 1744 counties in 37 states

28 states, 933 counties = N
36 states, 811 counties = Y

```{r}
d14 <- d %>%
  group_by(GEOID) %>%
  add_count(GEOID) %>%
  filter(n >= 14) %>% # Filter to >=14 corn yield observations
  ungroup(.) %>%
   dplyr::select(-n)


aggregate(d14$state_ansi, by = list(d14$state_ansi), FUN = length)
aggregate(d14$GEOID, by = list(d14$GEOID), FUN = length)

#Meng's code
d.new.14 <- d %>%
  mutate(irrigated = case_when(GEOID %in% d.irrgiated.filter$GEOID ~ 'N',
                               TRUE ~ 'Y')) %>%
  group_by(GEOID) %>%
  add_count(GEOID) %>%
  filter(n >= 14) %>% # Filter to >=14 corn yield observations #ATR: This is where he resricts the acres to 14 years of corn
  ungroup(.) %>%
  dplyr::select(-n)

d.new.N14 <-subset(d.new.14, irrigated == 'N')
d.new.Y14 <-subset(d.new.14, irrigated == 'Y')
aggregate(d.new.N14$state_ansi, by = list(d.new.N14$state_ansi, d.new.N14$irrigated ), FUN = length)
aggregate(d.new.N14$GEOID, by = list(d.new.N14$GEOID, d.new.N14$irrigated), FUN = length)
aggregate(d.new.Y14$state_ansi, by = list(d.new.Y14$state_ansi, d.new.Y14$irrigated ), FUN = length)
aggregate(d.new.Y14$GEOID, by = list(d.new.Y14$GEOID, d.new.Y14$irrigated), FUN = length)
```

ATR: With n= 14 d now has 34773 observations with 1695 counties in 37 states
28 states, 909 counties = N
36 states, 786 counties = Y

```{r}
d15 <- d %>%
  group_by(GEOID) %>%
  add_count(GEOID) %>%
  filter(n >= 15) %>% # Filter to >=15 corn yield observations
  ungroup(.) %>%
   dplyr::select(-n)

aggregate(d15$GEOID, by = list(d15$GEOID), FUN = length)
aggregate(d15$state_ansi, by = list(d15$state_ansi), FUN = length)

#Meng's code
d.new.15 <- d %>%
  mutate(irrigated = case_when(GEOID %in% d.irrgiated.filter$GEOID ~ 'N',
                               TRUE ~ 'Y')) %>%
  group_by(GEOID) %>%
  add_count(GEOID) %>%
  filter(n >= 15) %>% # Filter to >=15 corn yield observations #ATR: This is where he resricts the acres to 15 years of corn
  ungroup(.) %>%
  dplyr::select(-n)

d.new.N15 <-subset(d.new.15, irrigated == 'N')
d.new.Y15 <-subset(d.new.15, irrigated == 'Y')
aggregate(d.new.N15$state_ansi, by = list(d.new.N15$state_ansi, d.new.N15$irrigated ), FUN = length)
aggregate(d.new.N15$GEOID, by = list(d.new.N15$GEOID, d.new.N15$irrigated), FUN = length)
aggregate(d.new.Y15$state_ansi, by = list(d.new.Y15$state_ansi, d.new.Y15$irrigated ), FUN = length)
aggregate(d.new.Y15$GEOID, by = list(d.new.Y15$GEOID, d.new.Y15$irrigated), FUN = length)
```

ATR: With n= 15 d now has 34003 observations with 1640 counties in 36 states
27 states, 881 counties = N
35 states, 759 counties = Y

```{r}
d16 <- d %>%
  group_by(GEOID) %>%
  add_count(GEOID) %>%
  filter(n >= 16) %>% # Filter to >=16 corn yield observations
  ungroup(.) %>%
   dplyr::select(-n)

aggregate(d16$GEOID, by = list(d16$GEOID), FUN = length)
aggregate(d16$state_ansi, by = list(d16$state_ansi), FUN = length)

#Meng's code
d.new.16 <- d %>%
  mutate(irrigated = case_when(GEOID %in% d.irrgiated.filter$GEOID ~ 'N',
                               TRUE ~ 'Y')) %>%
  group_by(GEOID) %>%
  add_count(GEOID) %>%
  filter(n >= 16) %>% # Filter to >=16 corn yield observations #ATR: This is where he resricts the acres to 16 years of corn
  ungroup(.) %>%
  dplyr::select(-n)

d.new.N16 <-subset(d.new.16, irrigated == 'N')
d.new.Y16 <-subset(d.new.16, irrigated == 'Y')
aggregate(d.new.N16$state_ansi, by = list(d.new.N16$state_ansi, d.new.N16$irrigated ), FUN = length)
aggregate(d.new.N16$GEOID, by = list(d.new.N16$GEOID, d.new.N16$irrigated), FUN = length)
aggregate(d.new.Y16$state_ansi, by = list(d.new.Y16$state_ansi, d.new.Y16$irrigated ), FUN = length)
aggregate(d.new.Y16$GEOID, by = list(d.new.Y16$GEOID, d.new.Y16$irrigated), FUN = length)
```

ATR: With n= 16 d now has 32878 observations with 1565 counties in 36 states
27 states, 841 counties = N
35 states, 724 counties = Y

```{r}
d17 <- d %>%
  group_by(GEOID) %>%
  add_count(GEOID) %>%
  filter(n >= 17) %>% # Filter to >=17 corn yield observations
  ungroup(.) %>%
   dplyr::select(-n)

aggregate(d17$GEOID, by = list(d17$GEOID), FUN = length)
aggregate(d17$state_ansi, by = list(d17$state_ansi), FUN = length)

#Meng's code
d.new.17 <- d %>%
  mutate(irrigated = case_when(GEOID %in% d.irrgiated.filter$GEOID ~ 'N',
                               TRUE ~ 'Y')) %>%
  group_by(GEOID) %>%
  add_count(GEOID) %>%
  filter(n >= 17) %>% # Filter to >=17 corn yield observations #ATR: This is where he resricts the acres to 17 years of corn
  ungroup(.) %>%
  dplyr::select(-n)

d.new.N17 <-subset(d.new.17, irrigated == 'N')
d.new.Y17 <-subset(d.new.17, irrigated == 'Y')
aggregate(d.new.N17$state_ansi, by = list(d.new.N17$state_ansi, d.new.N17$irrigated ), FUN = length)
aggregate(d.new.N17$GEOID, by = list(d.new.N17$GEOID, d.new.N17$irrigated), FUN = length)
aggregate(d.new.Y17$state_ansi, by = list(d.new.Y17$state_ansi, d.new.Y17$irrigated ), FUN = length)
aggregate(d.new.Y17$GEOID, by = list(d.new.Y17$GEOID, d.new.Y17$irrigated), FUN = length)
```

ATR: With n= 17 d now has 31854 observations with 1501 counties in 36 states
27 states, 808 counties = N
35 states, 693 counties = Y

```{r}
d18 <- d %>%
  group_by(GEOID) %>%
  add_count(GEOID) %>%
  filter(n >= 18) %>% # Filter to >=18 corn yield observations
  ungroup(.) %>%
   dplyr::select(-n)

aggregate(d18$GEOID, by = list(d18$GEOID), FUN = length)
aggregate(d18$state_ansi, by = list(d18$state_ansi), FUN = length)

#Meng's code
d.new.18 <- d %>%
  mutate(irrigated = case_when(GEOID %in% d.irrgiated.filter$GEOID ~ 'N',
                               TRUE ~ 'Y')) %>%
  group_by(GEOID) %>%
  add_count(GEOID) %>%
  filter(n >= 18) %>% # Filter to >=18 corn yield observations #ATR: This is where he resricts the acres to 18 years of corn
  ungroup(.) %>%
  dplyr::select(-n)

d.new.N18 <-subset(d.new.18, irrigated == 'N')
d.new.Y18 <-subset(d.new.18, irrigated == 'Y')
aggregate(d.new.N18$state_ansi, by = list(d.new.N18$state_ansi, d.new.N18$irrigated ), FUN = length)
aggregate(d.new.N18$GEOID, by = list(d.new.N18$GEOID, d.new.N18$irrigated), FUN = length)
aggregate(d.new.Y18$state_ansi, by = list(d.new.Y18$state_ansi, d.new.Y18$irrigated ), FUN = length)
aggregate(d.new.Y18$GEOID, by = list(d.new.Y18$GEOID, d.new.Y18$irrigated), FUN = length)
```

ATR: With n= 18 d now has 30205 observations with 1404 counties in 31 states

Here is where I stopped. In 5 years we lost 340 counties and 6 states

```{r}
df <- data.frame(N = c(9, 10, 11, 12, 13, 14, 15, 16, 17, 18),
                 States = c(41, 40, 37, 37, 37, 37, 36, 36, 36, 31),
                 N_States = c(29, 29, 28, 28, 28, 28, 27, 27, 27, 25),
                 Y_States = c(40, 39, 36, 36, 36, 36, 35, 35, 35, 31), 
                 Counties = c(1914, 1858, 1818, 1782, 1744, 1695, 1640, 1565, 1501, 1404),
                 N_Counties = c(1012, 985, 969, 951, 933, 909, 881, 841, 808, 767), 
                 Y_Counties = c(902, 873, 849, 831, 811, 786, 759, 724, 693, 637)
                 )
df
```


ATR: Stopped here, waiting to hear which data set we use. 


```{r}
mod <- function(df){
  df <- df
  
  grid <- expand.grid(span = seq(0.3, 0.5, len = 5), degree = seq(0,1, len=2) )
  
  grid.control <- trainControl(
    method = "repeatedcv",
    number = 10,
    repeats = 5,
    search = "grid")
  
  train_loess <- train(Yield_mg_ha ~ year, 
                       method = "gamLoess",
                       tuneGrid=grid,
                       trControl=grid.control,
                       data = df)
  
  df$Detrend_resids <- as.numeric(residuals(train_loess))
  df$Detrend_predictions <- as.numeric(predict(train_loess))
  return(df)
}

```

ATR: Tried running it with d18 but didn't get far. 


```{r}
d18$year <- as.integer(d18$year) # year needs to be converted to integer to run the regression model in the train function above

# Dan's code using mclapply function
d_list <- split(d18, f = d18$GEOID) # ML: it creates a list based on county
```


```{r}
#test_list <- d_list[1:2]

tic()
d_list <- mclapply(X = d_list,FUN = mod, mc.cores = 1) 
# ML:seems like only one core can be used for windows 
toc()
# 4077.25 sec elapsed
```


```{r}
d.1 <- dplyr::bind_rows(d_list)

d.2 <- d.1 %>%
  group_by(GEOID) %>%
  mutate(County_avg_yield = mean(Yield_mg_ha)) %>%
  ungroup(.) %>%
  mutate(Yield_decomp_add = County_avg_yield+Detrend_resids,  # de-trended yield 
         Yield_decomp_mult = Yield_mg_ha/Detrend_predictions) # yield anomaly

```



```{r}


# saved the data to a rds file to be used for downstream processing
library(readr)
write_rds(d.2, file = "data/corn_yield_2000-2022_wo_irrigation.rds")


```

```{r}
# ML: didn't run
# ML: I also tried to use the purrr:map function to replicate the above mclappy function, and it was very slow; I then used furrr:future_map. This function is doing the same thing as purrr::map, but allows for running the script in parallel. 

library(purrr)
library(furrr) # this will allow the parallel computing of the purrr functions

tic()
data_list <- d17 %>%
  group_by(GEOID) %>%
  nest() %>%
  mutate(data = future_map(data, .f = mod))
toc()
# even with parallel, this step is still very slow; probably better to run through server

```

