---
title: "Untitled"
output: html_document
date: "2023-05-04"
---


# setup
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(rnassqs)
library(dplyr)
library(ggplot2)
library(tidyr)
library(reshape2)
library(caret)
library(parallel)
library(readr)
library(purrr)
api_key <- as.character(read.csv("/home/shared/Kane_data/nass_key.csv", header = F)[1,1])       # api key

# Call in all corn yield data via NASS API ####

nassqs_auth(key = api_key)

```

# Pull data from NASS   ##### OR YOU CAN SKIP THIS AND GO TO LINE 200 "d.1 <- read_rds(...)" ##### as of 3/29/23

1. Specify years and pull yield data from nass 'SURVEY'

```{r yield}
# Specify the range of years across which you want to collect data
years <- as.list(2000:2022)  
# ML: changed to 2000:2022 to include more data points
# 2000-2022: 38847 obs
# 2000-2016: 30108 obs

# BM: Future self, when update with newer data: consider which years to use to minimize 
# the confounding effect of improvements in maize genetics, i.e., shift the
# 16 year window forward, drop oldest records.

## Yields

d.soy <- plyr::ldply(years, function(x){
  
  params <- list(
    commodity_desc = "SOYBEANS",
    util_practice_desc = "ALL UTILIZATION PRACTICES",
    year = x,
    agg_level_desc = "COUNTY", 
    source_desc = "SURVEY",  # change source to source_desc
    domain_desc = "TOTAL"
  )
  
  return(
    nassqs_yields(params) %>%
      filter(
        prodn_practice_desc == "ALL PRODUCTION PRACTICES",
        county_ansi != ""
      ) %>%
      mutate(
        GEOID = paste(state_ansi, county_ansi, sep = ""),
        Yield_mg_ha = as.numeric(Value) * 0.0628
      ) %>%
      dplyr::select(
        year,
        GEOID,
        state_alpha,
        state_ansi,
        county_ansi,
        county_name, 
        Yield_mg_ha
        )
  )
})

#ATR addition: write the pulled nass data as it's own data file so we no longer have to pull it every time
#write_rds(d.soy, path = "/home/aysha/Kane_data/nass_soybean_050423.rds")
```

```{r}
d.soy <-read_rds("/home/aysha/Kane_data/nass_soybean_050423.rds")
```

2. Calculate irrigation acres from census data and create a filter
Note: we can skip this step if we'd like to keep data from irrigated land (Aysha)
Meng: I added the 2017 new census year

```{r irrgation}
### Total acres
census.years <- as.list(c(1997,2002,2007,2012, 2017))

d.acres.total.soy <- plyr::ldply(census.years, function(x) {
  
  params <- list(
    commodity_desc = "SOYBEANS",
    util_practice_desc = "ALL UTILIZATION PRACTICES",
    source_desc = "CENSUS",
    year = x,
    agg_level_desc = "COUNTY",
    short_desc = "SOYBEANS - ACRES HARVESTED",
    domain_desc = "TOTAL"
  )
  
  return(
    nassqs(params) %>%
      filter(county_ansi != "") %>%
      mutate(
        GEOID = paste(state_ansi, county_ansi, sep = ""),
        Acres_total = as.numeric(gsub(
          Value, pattern = ",", replacement = ""
        ))
      ) %>%
      dplyr::select(
        year,
        GEOID,
        state_alpha,
        state_ansi,
        county_ansi,
        county_name,
        Acres_total
      )
  )
})


```

```{r}
##### IRRIGATED ACRES

d.acres.soy.irrigated <- plyr::ldply(census.years, function(x) {
  
  params <- list(
    commodity_desc = "SOYBEANS",
    util_practice_desc = "ALL UTILIZATION PRACTICES",
    source_desc = "CENSUS",
    year = x,
    agg_level_desc = "COUNTY",
    short_desc = "SOYBEANS, IRRIGATED - ACRES HARVESTED",
    domain_desc = "TOTAL"
  )
  
  return(
    nassqs(params) %>%
      filter(county_ansi != "") %>%
      mutate(
        GEOID = paste(state_ansi, county_ansi, sep = ""),
        Acres_irrigated = as.numeric(gsub(
          Value, pattern = ",", replacement = ""
        ))
      ) %>%
      dplyr::select(
        year,
        GEOID,
        state_alpha,
        state_ansi,
        county_ansi,
        county_name,
        Acres_irrigated
      )
  )
})




```


```{r}
##

d.acres.soy <- d.acres.total.soy %>%
  left_join(d.acres.soy.irrigated) %>%
  filter(GEOID %in% d.soy$GEOID,!is.na(Acres_total)) %>%
  replace_na(list(Acres_irrigated = 0)) %>%
  mutate(Percent_irrigated = Acres_irrigated / Acres_total) %>%
  group_by(GEOID) %>%
  summarize(
    Mean.percent.irrigated = mean(Percent_irrigated),
    SD.percent.irrigated = sd(Percent_irrigated)
  )

## FILTER BASED ON IRRIGATED ACRES DATA

# Create filter to select counties that are 5 percent or less irrigated, 
# choice of 5 percent based on dsitribution of percentages, vast majority of counties are 5 percent or less irrigated

d.irrgiated.filter.soy <- d.acres.soy %>%
  filter(Mean.percent.irrigated <= 0.05) %>%
  filter(SD.percent.irrigated <= 0.01) 
```


3. filter yield data based on the availability of year numbers
ML: Instead of removing data from irrigated land, a new variable 'irrigated' was created:
irrigated == 'irrigated' means more than 95% of acres of a county were irrigated
irrigated == 'rainfed' means less than 5% of acres of a county were irrigated

```{r filter}

#Ran for 15 year filter 


d.soy.15 <- d.soy %>%
  mutate(irrigated = case_when(GEOID %in% d.irrgiated.filter.soy$GEOID ~ 'rainfed',
                               TRUE ~ 'irrigated')) %>%
  group_by(GEOID) %>%
  add_count(GEOID) %>%
  filter(n >= 15) %>% # Filter to >=15 corn yield observations #ATR: This is where he resricts the acres to 15 years of corn
  ungroup(.) %>%
  dplyr::select(-n) #18958 obs

```

4. extract de-trended yield

Note: used gamloess method to model the relationship between yield and year. 
The model was run for different counties

```{r detrend}
#install.packages("tictoc")
library(tictoc) # this is to record the running time

mod <- function(df){
  df <- df
  
  grid <- expand.grid(span = seq(0.3, 0.5, len = 5), degree = seq(0,1, len=2) )
  
  grid.control <- trainControl(
    method = "repeatedcv",
    number = 10,
    repeats = 5,
    search = "grid")

  train_loess <- caret::train(Yield_mg_ha ~ year, 
                       method = "gamLoess",
                       tuneGrid=grid,
                       trControl=grid.control,
                       data = df)
  
  df$Detrend_resids <- as.numeric(residuals(train_loess))
  df$Detrend_predictions <- as.numeric(predict(train_loess))
  return(df)
}



#---------------n = 22---------------#
d.soy.15$year <- as.integer(d.soy.15$year) # year needs to be converted to integer to run the regression model in the train function above


# Dan's code using mclapply function
d_soy_list_15 <- split(d.soy.15, f = d.soy.15$GEOID) # ML: it creates a list based on county

tic()
d_soy_list_15 <- mclapply(X = d_soy_list_15,FUN = mod, mc.cores = 28) 
toc()
# 161.737 sec elapsed

d.soy.15 <- dplyr::bind_rows(d_soy_list_15)

d.soy.15.1 <- d.soy.15 %>%
  group_by(GEOID) %>%
  mutate(County_avg_yield = mean(Yield_mg_ha)) %>%
  ungroup(.) %>%
  mutate(Yield_decomp_add = County_avg_yield+Detrend_resids,  # de-trended yield 
         Yield_decomp_mult = Yield_mg_ha/Detrend_predictions) # yield anomaly

# saved the data to a rds file to be used for downstream processing
write_rds(d.soy.15.1, file = "/home/aysha/Kane_data/soy_yield_2000-2022_w_irrigation_n_15.rds")

```


